## Qualitative content analysis

Each row represents one evaluative meaning unit from A305. Participant wording is retained as closely as possible; the theme and intensity are researcher-assigned codes. Use the search box or column filters to explore the table.

[Download the coded content analysis as an Excel workbook](../data/a305_content_analysis.xlsx)

```{r a305-content-analysis}
#| echo: false

a305_content_analysis <- readxl::read_xlsx(
  project_path("data", "a305_content_analysis.xlsx"),
  sheet = "A305 content analysis",
  col_types = "text"
)

reactable::reactable(
  a305_content_analysis,
  searchable = TRUE,
  filterable = TRUE,
  striped = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(10, 20, 50, 100),
  columns = list(
    Visualization = reactable::colDef(minWidth = 145),
    `Meaning unit` = reactable::colDef(minWidth = 420),
    Theme = reactable::colDef(minWidth = 160),
    `Evaluation intensity` = reactable::colDef(
      minWidth = 150,
      align = "center"
    )
  )
)
```

## Original responses

```{r}
feedback_A305 <- data %>%
  transmute(
    ID,
    feedback = stringr::str_squish(as.character(A305_01))
  ) %>%
  mutate(
    feedback = na_if(feedback, ""),
    feedback = na_if(feedback, "-9"),
    feedback = na_if(feedback, "NA")
  ) %>%
  filter(!is.na(feedback))
```

```{r}
feedback_A306 <- data %>%
  transmute(
    ID,
    feedback = stringr::str_squish(as.character(A306_01))
  ) %>%
  mutate(
    feedback = na_if(feedback, ""),
    feedback = na_if(feedback, "-9"),
    feedback = na_if(feedback, "NA")
  ) %>%
  filter(!is.na(feedback))
```

Comments specifically about the visualizations used in the study.

::: {.callout-note collapse="true"}
## Visualization feedback — Expand

```{r}
feedback_A305 %>%
  mutate(
    feedback = stringr::str_wrap(feedback, width = 100)
  ) %>%
  split(.$ID) %>%
  purrr::iwalk(function(x, id) {
    cat(
      "\nID:", id, "\n",
      x$feedback,
      "\n",
      paste(rep("-", 100), collapse = ""),
      "\n",
      sep = ""
    )
  })
```

:::

Broader comments about the study, tasks, or participants' experience.

::: {.callout-note collapse="true"}
## General feedback — Expand

```{r}
feedback_A306 %>%
  mutate(
    feedback = stringr::str_wrap(feedback, width = 100)
  ) %>%
  split(.$ID) %>%
  purrr::iwalk(function(x, id) {
    cat(
      "\nID:", id, "\n",
      x$feedback,
      "\n",
      paste(rep("-", 100), collapse = ""),
      "\n",
      sep = ""
    )
  })
```

:::

## Optional feedback export to PDF

```{r}
feedback_to_pdf <- function(df, title, filename) {

  df <- df %>%
    filter(!is.na(feedback), feedback != "") %>%
    arrange(ID)

  comments <- paste0(
    "<section class='comment'>",
    "<div class='id'>ID: ",
    htmltools::htmlEscape(as.character(df$ID)),
    "</div>",
    "<div class='text'>",
    htmltools::htmlEscape(df$feedback),
    "</div>",
    "</section>"
  )

  html <- paste0(
    "<!DOCTYPE html>
    <html>
    <head>
    <meta charset='UTF-8'>
    <style>
      @page {
        size: A4;
        margin: 18mm;
      }

      body {
        font-family: Arial, sans-serif;
        font-size: 11pt;
        line-height: 1.45;
        color: #222;
      }

      h1 {
        font-size: 18pt;
        margin-bottom: 20px;
      }

      .comment {
        padding: 10px 0 14px 0;
        border-bottom: 1px solid #cccccc;
        page-break-inside: avoid;
      }

      .id {
        font-weight: bold;
        margin-bottom: 5px;
      }

      .text {
        white-space: pre-wrap;
        overflow-wrap: anywhere;
      }
    </style>
    </head>

    <body>
      <h1>", htmltools::htmlEscape(title), "</h1>",
      paste(comments, collapse = "\n"),
    "</body>
    </html>"
  )

  html_file <- tempfile(fileext = ".html")
  writeLines(html, html_file, useBytes = TRUE)

  output_file <- normalizePath(filename, mustWork = FALSE)

  pagedown::chrome_print(
    input = html_file,
    output = output_file
  )

  message("Saved to: ", output_file)
}
```

```{r}
if (isTRUE(params$export_artifacts)) {
  feedback_to_pdf(
    feedback_A305,
    title = "Feedback - A305",
    filename = "feedback_A305.pdf"
  )

  feedback_to_pdf(
    feedback_A306,
    title = "Feedback - A306",
    filename = "feedback_A306.pdf"
  )
}
```
