---
title: "Analysis"
format:
  html:
    toc: true
    code-fold: true
    grid:
      body-width: 1200px
execute:
  warning: false
  message: false
params:
  export_artifacts: true
---

# Setup and data preparation

```{r setup}
library(dplyr)
library(ggplot2)
library(forcats)
library(tidyr)
library(glmmTMB)
library(emmeans)
library(ordinal)
library(purrr)
library(lme4)
library(patchwork)

sysfonts::font_add_google("Lato", family = "Lato")
showtext::showtext_auto()

vistypes <- c(
  "adjacent", "animation", "bivariate_corner", "bivariate_range",
  "triple_CI", "uncert_markers",  "basic"
)
vistype_labels <- c(
  adjacent = "Adjacent plots",
  animation = "Animation",
  bivariate_corner = "Bivariate corner",
  bivariate_range = "Bivariate range",
  triple_CI = "Confidence\nintervals",
  uncert_markers = "Uncertainty\nmarkers",
  basic = "Basic topoplot"
)

plot_theme <- function(base_size = 11) {
  theme_classic(base_size = base_size) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
      axis.ticks.y = element_blank()
    )
}

remove_group_iqr_outliers <- function(df, value, group) {
  value <- rlang::enquo(value)
  group <- rlang::enquo(group)

  df %>%
    group_by(!!group) %>%
    mutate(
      .q1 = quantile(!!value, 0.25, na.rm = TRUE),
      .q3 = quantile(!!value, 0.75, na.rm = TRUE),
      .iqr = .q3 - .q1
    ) %>%
    filter(
      !!value >= .q1 - 1.5 * .iqr,
      !!value <= .q3 + 1.5 * .iqr
    ) %>%
    ungroup() %>%
    select(-.q1, -.q3, -.iqr)
}
```

## Import and participant inclusion

```{r import-data}
survey_raw <- read.csv2(
  "data_uncertain_topoplots.csv",
  fileEncoding = "UTF-16LE",
  check.names = FALSE
)

legend <- survey_raw %>% dplyr::slice(1)

required_columns <- c("STARTED", "TIME051", "TIME_SUM", "A001")
missing_columns <- setdiff(required_columns, names(survey_raw))
if (length(missing_columns) > 0) {
  stop("Missing required columns: ", paste(missing_columns, collapse = ", "))
}

data_full <- survey_raw
data <- survey_raw %>%
  dplyr::slice(-1) %>%
  filter(as.Date(STARTED) >= as.Date("2026-07-24")) %>%
  filter(!is.na(suppressWarnings(as.numeric(TIME051)))) %>%
  select(-(1:5)) %>%
  mutate(ID = row_number(), .before = 1)
```

```{r sample-summary}
tibble(
  respondents = nrow(data),
  mean_completion_minutes = mean(as.numeric(data$TIME_SUM), na.rm = TRUE) / 60,
  median_completion_minutes = median(as.numeric(data$TIME_SUM), na.rm = TRUE) / 60
)
```

# Participant characteristics

## Self-assessed visualization experience
```{r}
n_total <- data %>%
  filter(!is.na(A001)) %>%
  nrow()

data$A001 %>% table() %>%  data.frame() %>% dplyr::rename(level = !!names(.)[1]) %>%
    mutate(percent_score = round(Freq / sum(Freq) * 100)) %>% 
    mutate(level = factor(level, levels = rev(c("Beginner", "Intermediate", "Expert")))) %>% 
    ggplot(data = ., aes(y = level, x = percent_score)) +  
    geom_col(stat="identity",  width = 0.8, fill ="#6BAED6") + labs(x = "Percent of respondents", y = "", title="Self-assessed level") +
    geom_text(aes(label = paste0(percent_score, "%")), hjust = -0.2)  +  theme_classic(base_family = "Lato") + 
 scale_x_continuous(
    name = "Percent of respondents",
    labels = \(x) paste0(x, "%"),
    breaks = seq(0, 50, 10),

    sec.axis = sec_axis(
      transform = ~ . * n_total / 100,
      name = "Number of respondents"
    ),

    expand = expansion(mult = c(0, 0.12))
  ) + theme(
    plot.title = element_text(
      size = 16,
      face = "bold", hjust = 0.5,
      margin = margin(b = 14)
    ),
    axis.title.x = element_text(
      size = 14, hjust = 0.5,
      margin = margin(t = 10),
    ),
    axis.text = element_text(size = 11),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )  + coord_cartesian(clip = "off")
```

## Country of work
```{r}
country <- data %>%
  transmute(
    country_work = case_when(
      A004 == "-2" & !is.na(A004s) & A004s != "" ~ A004s,
      TRUE ~ A004
    )
  ) %>% mutate(
    country_work = trimws(country_work),

    country_work = case_when(
      country_work %in% c("NL", "The Netherlands") ~ "Netherlands",
      country_work %in% c("United States", "United States of America") ~ "USA",
      country_work %in% c(
        "Prefer not to say",
        "Professor",
        "Senior Researcher"
      ) ~ NA_character_,
      TRUE ~ country_work
    )) %>%
  filter(!is.na(country_work), country_work != "") %>%
  mutate(
    country_work = fct_lump_min(
      factor(country_work),
      min = 2,
      other_level = "Other"
    )
  ) %>% count(country_work, name = "Freq") %>%
  mutate(
    percent_score = round(Freq / sum(Freq) * 100),
    country_work = fct_reorder(
      country_work,
      percent_score
    )
  )
country

```

### Diagnostic: responses grouped as “Other”

```{r country-other-diagnostic}
country_other <- data %>%
  transmute(
    country_original = case_when(
      A004 == "-2" & !is.na(A004s) & A004s != "" ~ A004s,
      TRUE ~ A004
    )
  ) %>%
  mutate(
    country_original = trimws(country_original),
    country_original = case_when(
      country_original %in% c("NL", "The Netherlands") ~ "Netherlands",
      country_original %in% c("United States", "United States of America") ~ "USA",
      country_original %in% c(
        "Prefer not to say",
        "Professor",
        "Senior Researcher"
      ) ~ NA_character_,
      TRUE ~ country_original
    )
  ) %>%
  filter(!is.na(country_original), country_original != "") %>%
  count(country_original, name = "respondents", sort = TRUE) %>%
  filter(respondents < 2)

country_other
```

```{r}

country %>%
  ggplot(aes(percent_score, reorder(country_work, percent_score))) +
  geom_col(width = 0.7, fill = "#6BAED6") +
  geom_text(aes(label = paste0(round(percent_score), "%")),
            hjust = -0.15, size = 4) +
  scale_x_continuous(
    "Percent of respondents",
    labels = \(x) paste0(x, "%"),
    sec.axis = sec_axis(~ . * n_total / 100, name = "Number of respondents"),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Country of work",
    y = NULL,
   # caption = paste("N of respondents =", n_total)
  ) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text = element_text(size = 12),
    axis.ticks.y = element_blank(),
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0)
  )
```

## Research field
```{r}
field <- data %>%
  mutate(field = stringr::word(A005, 1)) %>%
  count(field, name = "Freq") %>%
  mutate(percent_score = Freq / sum(Freq) * 100)
field
```
```{r}
field %>%
  ggplot(aes(percent_score, reorder(field, percent_score))) +
  geom_col(width = 0.7, fill = "#6BAED6") +
  geom_text(aes(label = paste0(round(percent_score), "%")),
            hjust = -0.15, size = 4) +
  scale_x_continuous(
    "Percent of respondents",
    labels = \(x) paste0(x, "%"),
    sec.axis = sec_axis(~ . * n_total / 100,
                        name = "Number of respondents"),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(title = "Research field", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text = element_text(size = 12),
    axis.ticks.y = element_blank()
  )
```
## Software used

```{r}
software_names <- legend %>%
  select(matches("^A006_\\d{2}$")) %>%
  pivot_longer(everything(), names_to = "item", values_to = "software")  %>%
  mutate(
    software = stringr::word(software, 2, -1),
    software = if_else(item == "A006_27", "Other", software)
  )


software <- data %>%
  select(matches("^A006_\\d{2}$")) %>%
  pivot_longer(everything(), names_to = "item", values_to = "response") %>%
  filter(response == "Checked") %>%
  count(item, name = "Freq") %>%
  right_join(software_names, by = "item") %>%
  mutate(
    Freq = coalesce(Freq, 0L),
    percent_score = Freq / n_total * 100
  )
software
```
```{r}
software %>% filter(Freq != 0) %>% 
  ggplot(aes(percent_score, reorder(software, percent_score))) +
  geom_col(width = 0.7, fill = "#6BAED6") +
  geom_text(aes(label = paste0(round(percent_score), "%")),
            hjust = -0.15, size = 4) +
  scale_x_continuous(
    "Percent of respondents",
    labels = \(x) paste0(x, "%"),
    sec.axis = sec_axis(~ . * n_total / 100,
                        name = "Number of respondents"),
    expand = expansion(mult = c(0, 0.18))
  ) +
  labs(title = "Software used", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text = element_text(size = 11),
    axis.ticks.y = element_blank()
  )
```

## Combined participant-characteristics figure

```{r participant-characteristics-figure}
#| fig-width: 15
#| fig-height: 13

format_category_labels <- function(labels) {
  lapply(
    labels,
    \(label) {
      if (identical(label, "Other")) {
        bquote(italic(.(label)))
      } else {
        label
      }
    }
  )
}

make_characteristic_plot <- function(
  plot_data,
  label_column,
  title,
  bar_width = 0.7
) {
  plot_data <- plot_data %>%
    filter(!is.na(.data[[label_column]])) %>%
    mutate(
      display_label = forcats::fct_reorder(
        factor(.data[[label_column]]),
        percent_score
      )
    )

  ggplot(plot_data, aes(x = percent_score, y = display_label)) +
    geom_col(width = bar_width, fill = "#6BAED6") +
    geom_text(
      aes(label = paste0(round(percent_score), "%")),
      hjust = -0.15,
      size = 8
    ) +
    scale_x_continuous(
      limits = c(0, 100),
      breaks = seq(0, 100, 25),
      labels = \(x) paste0(x, "%"),
      expand = expansion(mult = c(0, 0.02))
    ) +
    scale_y_discrete(labels = format_category_labels) +
    labs(title = title, x = "Percent of respondents", y = NULL) +
    coord_cartesian(clip = "off") +
    theme_classic(base_family = "Lato", base_size = 14) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
      axis.title.x = element_text(size = 15),
      axis.text.x = element_text(size = 13),
      axis.text.y = element_text(size = 20),
      axis.ticks.y = element_blank(),
      plot.margin = margin(8, 20, 8, 8)
    )
}

experience_combined <- data %>%
  filter(!is.na(A001), A001 != "") %>%
  count(level = A001, name = "Freq") %>%
  mutate(percent_score = Freq / sum(Freq) * 100)

country_combined <- country %>%
  transmute(
    country_work = as.character(country_work),
    Freq,
    percent_score
  )

field_combined <- field %>%
  filter(!is.na(field), field != "")

software_combined <- software %>%
  filter(Freq > 0, !is.na(software), software != "")

p_experience <- make_characteristic_plot(
  experience_combined,
  "level",
  "Self-assessed level"
)
p_country <- make_characteristic_plot(
  country_combined,
  "country_work",
  "Country of work"
)
p_field <- make_characteristic_plot(
  field_combined,
  "field",
  "Research field"
)
p_software <- make_characteristic_plot(
  software_combined,
  "software",
  "Software used"
)

participant_characteristics_figure <-
  (p_experience + p_field) /
  (p_country + p_software) +
  plot_layout(heights = c(0.65, 1.35)) +
  plot_annotation(
    tag_levels = "A",
    caption = paste0(
      "Bars show percentages of respondents with valid responses. ",
      "Multiple software selections were permitted."
    ),
    theme = theme(
      plot.tag = element_text(face = "bold", size = 18),
      plot.caption = element_text(
        hjust = 0.5,
        face = "italic",
        family = "Lato",
        size = 13
      )
    )
  )

participant_characteristics_figure

if (isTRUE(params$export_artifacts)) {
  ggsave(
    "participant_characteristics_figure.png",
    participant_characteristics_figure,
    width = 15,
    height = 13,
    dpi = 140
  )
}
```

### Reduced figure excluding categories at or below 3%

```{r participant-characteristics-figure-over-3-percent}
#| fig-width: 15
#| fig-height: 9

country_over_3 <- country_combined %>%
  mutate(percent_score = Freq / sum(Freq) * 100) %>%
  filter(round(percent_score) > 3)

software_over_3 <- software_combined %>%
  filter(round(percent_score) > 3)

p_country_over_3 <- make_characteristic_plot(
  country_over_3,
  "country_work",
  "Country of work"
)

p_software_over_3 <- make_characteristic_plot(
  software_over_3,
  "software",
  "Software used"
)

participant_characteristics_over_3_figure <-
  (p_experience + p_field) /
  (p_country_over_3 + p_software_over_3) +
  plot_layout(heights = c(0.65, 1.35)) +
  plot_annotation(
    tag_levels = "A",
    caption = paste0(
      "Categories ≥ 3% are combined into 'Other' in panels C and D. ",
      "Multiple software selections were permitted."
    ),
    theme = theme(
      plot.tag = element_text(face = "bold", size = 18),
      plot.caption = element_text(
        hjust = 0.5,
        face = "italic",
        family = "Lato",
        size = 13
      )
    )
  )

participant_characteristics_over_3_figure

if (isTRUE(params$export_artifacts)) {
  ggsave(
    "participant_characteristics_over_3_figure.png",
    participant_characteristics_over_3_figure,
    width = 15,
    height = 13,
    dpi = 140
  )
}

```

# Task 1: Visualization comparison

## Trial reconstruction and presentation order
```{r}
get_field <- function(x, field) {
  stringr::str_match(
    x,
    paste0("(?:^|\\|)", field, "=([^|\\n]*)")
  )[, 2]
}

task1_order <- data %>%
  dplyr::select(ID, A125_02) %>%
  tidyr::separate_rows(A125_02, sep = "\n") %>%
  dplyr::filter(A125_02 != "") %>%
  dplyr::transmute(
    ID,
    position = as.integer(get_field(A125_02, "position")),
    trial    = as.integer(get_field(A125_02, "trial")),
    question = get_field(A125_02, "question"),
    stimulus = get_field(A125_02, "stimulus"),
    expected = as.integer(get_field(A125_02, "expected")),
    answered = as.integer(get_field(A125_02, "answered")),
    correct  = as.integer(get_field(A125_02, "correct"))
  ) %>%
  dplyr::arrange(ID, position)

task1_design <- data %>%
  dplyr::select(ID, A124_01) %>%
  tidyr::separate_rows(A124_01, sep = "\n") %>%
  dplyr::filter(A124_01 != "") %>%
  dplyr::transmute(
    ID,
    trial       = as.integer(get_field(A124_01, "trial")),
    question    = get_field(A124_01, "question"),
    vistype     = get_field(A124_01, "vistype"),
    target      = get_field(A124_01, "target"),
    correct     = get_field(A124_01, "correct"),
    wrong       = get_field(A124_01, "wrong"),
    pair        = get_field(A124_01, "pair"),
    solution    = as.integer(get_field(A124_01, "solution")),
    opt1        = get_field(A124_01, "opt1"),
    opt2        = get_field(A124_01, "opt2")
  )

task1 <- task1_order %>%
  left_join(
    task1_design %>%
      select(ID, trial, question, vistype),
    by = c("ID", "trial", "question")
  )
```

## Accuracy analysis
```{r}
task1_model <- task1 %>%
  transmute(
    ID = factor(ID),
    trial = factor(trial),
    vistype = relevel(factor(vistype), ref = "basic"),
    position = as.integer(position),
    correct = as.integer(correct)
  ) %>%
  filter(!is.na(correct), !is.na(position), !is.na(vistype)) %>%
  mutate(position_c = position - mean(position))
```

```{r}
vistype_table <- task1_model |>
  group_by(vistype) |>
  summarise(
    n = n(),
    n_correct = sum(correct == 1, na.rm = TRUE),
    n_incorrect = sum(correct == 0, na.rm = TRUE),
    accuracy = mean(correct, na.rm = TRUE),
    sd_accuracy = sd(correct, na.rm = TRUE),
    .groups = "drop"
  )
vistype_table
```
```{r}
vistype_table |>
  ggplot(aes(accuracy, reorder(vistype, accuracy))) +
  geom_col(
    width = 0.7,
    fill = "#6BAED6"
  ) +
  geom_errorbar(
    aes(
      xmin = pmax(accuracy - sd_accuracy, 0),
      xmax = pmin(accuracy + sd_accuracy, 1)
    ),
    orientation = "y",
    width = 0.14,
    linewidth = 0.7,
    position = position_nudge(y = 0.17)
  ) +
  geom_text(
    aes(label = sprintf("%.2f", accuracy)),
    hjust = -0.15,
    nudge_y = -0.20,
    size = 4
  ) +
  scale_x_continuous(
    "Mean accuracy",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    labels = scales::label_percent(),
    expand = expansion(mult = c(0, 0.10))
  ) +
  scale_y_discrete(
    labels = vistype_labels
  ) +
  labs(
    title = "Comparison task accuracy",
    y = NULL,
    caption = "Bars show mean accuracy; error bars indicate ±1 SD."
  ) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato") +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16
    ),
    axis.text = element_text(size = 11),
    axis.ticks.y = element_blank(),
    plot.caption = element_text(
      hjust = 0.5,
      face = "italic"
    )
  )
```

```{r}

# Logistic mixed-effects model
task1_fit <- glmer(
  correct ~ vistype + position_c + (1 | ID),
  data = task1_model,
  family = binomial
)

# Estimated probabilities and 95% confidence intervals
accuracy_estimates <- emmeans(
  task1_fit,
  ~ vistype,
  type = "response"
) |>
  confint() |>
  as.data.frame() |>
  transmute(
    vistype,
    accuracy = prob,
    lower_ci = asymp.LCL,
    upper_ci = asymp.UCL
  ) |>
  arrange(desc(accuracy)) |>
  mutate(
    vistype = factor(vistype, levels = rev(vistype))
  )

accuracy_estimates
```
```{r}
# Model-estimated probability of a correct response for each visualization type.
ggplot(
  accuracy_estimates,
  aes(
    x = accuracy,
    y = vistype
  )
) +
  geom_errorbar(
    aes(
      xmin = lower_ci,
      xmax = upper_ci
    ),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_point(size = 3.5) +
  geom_text(
    aes(label = sprintf("%.2f", accuracy)),
    nudge_y = 0.36,
    vjust = 0.5,
    size = 4
  ) +
  scale_y_discrete(
    expand = expansion(add = 0.5),
    labels = vistype_labels
  ) +
  scale_x_continuous(
  labels = function(x) sprintf("%.2f", x),
  limits = c(0.5, 1),
  breaks = seq(0.5, 1, 0.1)
) +
  labs(
    title = "Comparison task accuracy",
    x = "Model-estimated accuracy",
    y = NULL,
    caption = "Points show model-estimated probabilities; error bars indicate 95% confidence intervals."
  ) +
  theme_classic(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.caption = element_text(hjust = 0.5, face = "italic")
  )
```


```{r}
model_order <- glmmTMB(
  correct ~ vistype + position_c +
    (1 | ID),
  family = binomial,
  data = task1_model
)

summary(model_order)
```
```{r}
model_no_vistype <- glmmTMB(
  correct ~ position_c + (1 | ID),
  family = binomial,
  data = task1_model
)

model_vistype <- glmmTMB(
  correct ~ vistype + position_c + (1 | ID),
  family = binomial,
  data = task1_model
)

anova(model_no_vistype, model_vistype)
```
## Completion-time analysis

```{r}
task1_time <- data %>%
  select(ID, TIME005:TIME025) %>%
  pivot_longer(-ID, names_to = "time_var", values_to = "time_sec") %>%
  mutate(
    ID = as.character(ID),
    # TIME005:TIME025 correspond to experimental positions 1:21.
    position = as.integer(stringr::str_remove(time_var, "TIME")) - 4L,
    time_sec = readr::parse_double(as.character(time_sec))
  ) %>%
  left_join(
    task1 %>%
      transmute(
        ID = as.character(ID),
        position,
        trial,
        vistype
      ),
    by = c("ID", "position")
  ) %>%
  filter(!is.na(time_sec), !is.na(vistype))

task1_time
```

```{r}
time_table <- task1_time %>%
  remove_group_iqr_outliers(time_sec, vistype) %>%
  group_by(vistype) %>%
  summarise(
    n = n(),
    median_sec = median(time_sec, na.rm = TRUE),
    mean_sec = mean(time_sec, na.rm = TRUE),
    sd_sec = sd(time_sec, na.rm = TRUE),
    .groups = "drop"
  )

time_table |> arrange((median_sec))
```
```{r}
task1_time %>%
  remove_group_iqr_outliers(time_sec, vistype) %>%
  ggplot(aes(vistype, time_sec)) + 
  geom_boxplot(outlier.alpha = 0.3) +
  labs(
    title = "Time spent by visualization type",
    x = NULL,
    y = "Time spent on page (seconds)"
  ) +
  theme_classic(base_family = "Lato") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 35, hjust = 1)
  )
```

### Model-estimated completion time

```{r task1-time-model}
task1_time_model <- task1_time %>%
  transmute(
    ID = factor(ID),
    vistype = relevel(factor(vistype), ref = "basic"),
    position = as.integer(position),
    time_sec = as.numeric(time_sec)
  ) %>%
  filter(
    !is.na(ID),
    !is.na(vistype),
    !is.na(position),
    !is.na(time_sec),
    time_sec > 0
  ) %>%
  mutate(position_c = position - mean(position))

task1_time_fit <- lmer(
  log(time_sec) ~ vistype + position_c + (1 | ID),
  data = task1_time_model
)

time_estimates <- emmeans(
  task1_time_fit,
  ~ vistype,
  type = "response"
) %>%
  confint() %>%
  as.data.frame() %>%
  transmute(
    vistype,
    estimated_time = response,
    lower_ci = lower.CL,
    upper_ci = upper.CL
  ) %>%
  arrange(estimated_time) %>%
  mutate(vistype = factor(vistype, levels = vistype))

time_estimates
```

```{r task1-time-estimate-plot}
# Model-estimated time for each visualization type.
ggplot(
  time_estimates,
  aes(x = estimated_time, y = vistype)
) +
  geom_errorbar(
    aes(xmin = lower_ci, xmax = upper_ci),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_point(size = 3.5) +
  geom_text(
    aes(label = sprintf("%.1f s", estimated_time)),
    nudge_y = 0.36,
    vjust = 0.5,
    size = 4
  ) +
  scale_y_discrete(
    expand = expansion(add = 0.5),
    labels = vistype_labels
  ) +
  scale_x_continuous(
    labels = \(x) sprintf("%.0f s", x),
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  labs(
    title = "Comparison task completion time",
    x = "Model-estimated time (seconds)",
    y = NULL,
    caption = paste(
      "Points show back-transformed model estimates;",
      "error bars indicate 95% confidence intervals."
    )
  ) +
  theme_classic(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.caption = element_text(hjust = 0.5, face = "italic"),
    axis.ticks.y = element_blank()
  )
```

# Task 2: Pattern detection

## Response scoring and trial reconstruction

```{r}
task2_order <- data %>%
  select(ID, A207_01:A207_21) %>%
  pivot_longer(
    -ID,
    names_to = "trial",
    names_prefix = "A207_",
    values_to = "position"
  ) %>%
  mutate(
    ID = as.character(ID),
    trial = as.integer(trial),
    position = as.integer(position)
  )
```

```{r}
task2_design <- data %>%
  select(ID, A205_01) %>%
  tidyr::separate_rows(A205_01, sep = "###") %>%
  filter(A205_01 != "") %>%
  transmute(
    ID,
    trial   = as.integer(get_field(A205_01, "trial")),
    vistype = get_field(A205_01, "vistype"),
    pattern = get_field(A205_01, "pattern"),
    subject = as.integer(get_field(A205_01, "subject")),
    stimulus = get_field(A205_01, "stim"),
    img_url  = get_field(A205_01, "imgUrl"),
    input_id = get_field(A205_01, "inputId")
  )

task2_responses <- data %>%
  select(ID, A201_01:A201_21) %>%
  pivot_longer(
    -ID,
    names_to = "input_id",
    values_to = "selected_region"
  )
```


```{r}
score_text <- paste(
  readLines("regions.txt", encoding = "UTF-8"),
  collapse = "\n"
)
score_matches <- stringr::str_match_all(
  score_text,
  '"([^"]+)"\\s*=>\\s*array\\(([^)]*)\\)'
)[[1]]
region_scores <- tibble::tibble(
  stimulus = score_matches[, 2],
  values = score_matches[, 3]
) %>%
  tidyr::separate_rows(values, sep = ",\\s*") %>%
  group_by(stimulus) %>%
  mutate(
    region = sprintf("R%02d", row_number()),
    score = as.numeric(values)
  ) %>%
  ungroup() %>%
  select(stimulus, region, score)
```

```{r}
task2 <- task2_design %>%
  mutate(ID = as.character(ID)) %>%
  left_join(task2_responses %>% mutate(ID = as.character(ID)),
            by = c("ID", "input_id")) %>%
  mutate(
    region_number = readr::parse_number(selected_region),
    region = sprintf("R%02d", region_number)
  ) %>%
  left_join(region_scores, by = c("stimulus", "region")) %>%
  left_join(task2_order, by = c("ID", "trial")) %>%
  select(ID, trial, position, vistype, stimulus, region, score)

task2
```




## Descriptive detection scores

```{r}
task2_model <- task2 %>%
  transmute(
    ID = factor(ID),
    stimulus = factor(stimulus),
    vistype = relevel(factor(vistype), ref = "basic"),
    position_c = as.integer(position) - 11,
    score = as.numeric(score)
  ) %>%
  filter(complete.cases(.))
task2_model
```

```{r}
task2_model[task2_model$stimulus == "MMN_30",]
```

```{r task2-descriptive-summary}
task2_table <- task2_model %>%
  group_by(vistype) %>%
  summarise(
    n = n(),
    mean = round(mean(score), 3),
    median = round(median(score), 3),
    sd = round(sd(score), 3),
    .groups = "drop"
  )

task2_table |> arrange(desc(median))
```

```{r task2-descriptive-plot}
task2_table |>
  ggplot(aes(median, reorder(vistype, median))) +
  geom_col(width = 0.7, fill = "#6BAED6") +
  geom_errorbar(
    aes(
      xmin = pmax(median - sd, 0),
      xmax = pmin(median + sd, 1)
    ),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_text(
    aes(label = sprintf("%.2f", median)),
    hjust = -0.15,
    nudge_y = -0.20,
    size = 4
  ) +
  scale_x_continuous(
    "Median score",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.10))
  ) +
  scale_y_discrete(labels = vistype_labels) +
  labs(title = "Detection task accuracy", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text = element_text(size = 11),
    axis.ticks.y = element_blank()
  )
```

## Detection-score analysis

```{r}
model_task2 <- glmmTMB(
  score ~ vistype + 
    (1 | ID),
  family = ordbeta(link = "logit"),
  data = task2_model
)

summary(model_task2)
```

```{r}
model_task2_no_vistype <- update(
  model_task2,
  . ~ . - vistype
)

anova(model_task2_no_vistype, model_task2)
```

```{r task2-estimates}
task2_estimates <- emmeans(
  model_task2,
  ~ vistype,
  type = "response"
) |>
  confint() |>
  as.data.frame() |>
  transmute(
    vistype,
    predicted_score = response,
    lower_ci = asymp.LCL,
    upper_ci = asymp.UCL
  ) |>
  arrange(predicted_score) |>
  mutate(
    vistype = factor(vistype, levels = vistype)
  )

task2_estimates
```
```{r task2-pairwise-comparisons}
emm_task2 <- emmeans(
  model_task2,
  ~ vistype
)

task2_pairs <- pairs(
  emm_task2,
  adjust = "tukey"
)

task2_pairs

task2_pairs |>
  as.data.frame() |>
  dplyr::filter(p.value < 0.05)
```
```{r task2-response-scale-comparisons}
emm_task2_response <- regrid(
  emm_task2,
  transform = "response"
)

task2_score_differences <- pairs(
  emm_task2_response,
  adjust = "tukey"
) |>
  confint() |>
  as.data.frame()

task2_score_differences
```


```{r}

ggplot(
  task2_estimates,
  aes(
    x = predicted_score,
    y = vistype
  )
) +
  geom_errorbar(
    aes(
      xmin = lower_ci,
      xmax = upper_ci
    ),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_point(size = 3.5) +
  geom_text(
    aes(label = sprintf("%.2f", predicted_score)),
    nudge_y = 0.36,
    vjust = 0.5,
    size = 4
  ) +
  scale_y_discrete(
    labels = vistype_labels,
    expand = expansion(add = 0.5)
  ) +
  scale_x_continuous(
    labels = function(x) sprintf("%.2f", x),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2)
  ) +
  labs(
    title = "Detection task score",
    x = "Model-estimated score",
    y = NULL,
    caption = paste(
      "Points show model-estimated scores;",
      "error bars indicate 95% confidence intervals."
    )
  ) +
  theme_classic(base_family = "Lato", base_size = 15) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16
    ),
    plot.caption = element_text(
      hjust = 0.5,
      face = "italic"
    ),
    axis.text = element_text(size = 11),
    axis.ticks.y = element_blank()
  )
```

<!-- Descriptive code moved before the model analysis. -->
```{r}
#| eval: false
#| include: false
task2_table <- task2_model %>%
  group_by(vistype) %>%
  summarise(
    n = n(),
    mean = round(mean(score), 3),
    median = round(median(score), 3),
    sd = round(sd(score), 3),
    .groups = "drop"
  )

task2_table |> arrange(desc(median))
```
```{r}
#| eval: false
#| include: false
task2_table |>
  ggplot(aes(median, reorder(vistype, median))) +
  geom_col(
    width = 0.7,
    fill = "#6BAED6"
  ) +
  geom_errorbar(
    aes(
      xmin = pmax(median - sd, 0),
      xmax = pmin(median + sd, 1)
    ),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_text(
    aes(label = sprintf("%.2f", median)),
    hjust = -0.15,
    nudge_y = -0.20,
    size = 4
  ) +
  scale_x_continuous(
    "Median score",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.10))
  ) +
  scale_y_discrete(
    labels = c(
      adjacent = "Adjacent\nplots",
      animation = "Animation",
      bivariate_corner = "Bivariate\ncorner",
      bivariate_range = "Bivariate\nrange",
      triple_CI = "Confidence\nintervals",
      uncert_markers = "Uncertainty\nmarkers",
      basic = "Basic\ntopoplot"
    )
  ) +
  labs(
    title = "Detection task accuracy",
    y = NULL
  ) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato") +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 16
    ),
    axis.text = element_text(size = 11),
    axis.ticks.y = element_blank()
  )
```


## Completion-time analysis

```{r task2-time-data}
task2_time <- data %>%
  select(ID, TIME029:TIME049) %>%
  pivot_longer(
    -ID,
    names_to = "time_var",
    values_to = "time_sec"
  ) %>%
  mutate(
    ID = as.character(ID),
    # TIME029:TIME049 correspond to experimental positions 1:21.
    position = as.integer(stringr::str_remove(time_var, "TIME")) - 28L,
    time_sec = readr::parse_double(as.character(time_sec))
  ) %>%
  left_join(
    task2 %>%
      transmute(
        ID = as.character(ID),
        position,
        trial,
        vistype
      ),
    by = c("ID", "position")
  ) %>%
  filter(!is.na(time_sec), !is.na(vistype), time_sec > 0)

task2_time
```

```{r task2-time-summary}
task2_time_table <- task2_time %>%
  remove_group_iqr_outliers(time_sec, vistype) %>%
  group_by(vistype) %>%
  summarise(
    n = n(),
    median_sec = median(time_sec, na.rm = TRUE),
    mean_sec = mean(time_sec, na.rm = TRUE),
    sd_sec = sd(time_sec, na.rm = TRUE),
    .groups = "drop"
  )

task2_time_table |> arrange(median_sec)
```

### Model-estimated completion time

```{r task2-time-model}
task2_time_model <- task2_time %>%
  transmute(
    ID = factor(ID),
    vistype = relevel(factor(vistype), ref = "basic"),
    position = as.integer(position),
    time_sec = as.numeric(time_sec)
  ) %>%
  filter(complete.cases(.), time_sec > 0) %>%
  mutate(position_c = position - mean(position))

task2_time_fit <- lmer(
  log(time_sec) ~ vistype + position_c + (1 | ID),
  data = task2_time_model
)

task2_time_estimates <- emmeans(
  task2_time_fit,
  ~ vistype,
  type = "response"
) %>%
  confint() %>%
  as.data.frame() %>%
  transmute(
    vistype,
    estimated_time = response,
    lower_ci = lower.CL,
    upper_ci = upper.CL
  ) %>%
  arrange(estimated_time) %>%
  mutate(vistype = factor(vistype, levels = vistype))

task2_time_estimates
```

```{r task2-time-estimate-plot}
#| fig-width: 9
#| fig-height: 5.5
ggplot(
  task2_time_estimates,
  aes(x = estimated_time, y = vistype)
) +
  geom_errorbar(
    aes(xmin = lower_ci, xmax = upper_ci),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_point(size = 3.5) +
  geom_text(
    aes(label = sprintf("%.1f s", estimated_time)),
    nudge_y = 0.36,
    vjust = 0.5,
    size = 4
  ) +
  scale_y_discrete(
    expand = expansion(add = 0.5),
    labels = vistype_labels
  ) +
  scale_x_continuous(
    labels = \(x) sprintf("%.0f s", x),
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  labs(
    title = "Detection task completion time",
    x = "Model-estimated time (seconds)",
    y = NULL,
    caption = paste0(
      "Points show back-transformed model estimates.\n",
      "Error bars indicate 95% confidence intervals."
    )
  ) +
  theme_classic(base_size = 15) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.caption = element_text(hjust = 0.5, face = "italic"),
    axis.ticks.y = element_blank(),
    plot.margin = margin(8, 16, 12, 16)
  )
```

# Subjective evaluation

## Rating data preparation and descriptive summary

```{r}

question_labels <- c(
  A301 = "Helpful for comparison",
  A302 = "Helpful for detection and integration",
  A303 = "Helpful for representation",
  A304 = "Aesthetically appealing"
)

likert <- data %>%
  select(ID, matches("^A30[1-4]_\\d{2}$")) %>%
  pivot_longer(
    -ID,
    names_to = c("question_code", "number"),
    names_pattern = "(A30[1-4])_(\\d{2})",
    values_to = "rating_raw"
  ) %>%
  mutate(
    number = as.integer(number),

    vistype = factor(
      vistypes[number],
      levels = vistypes
    ),

    question = factor(
      question_code,
      levels = names(question_labels),
      labels = unname(question_labels)
    ),

    rating = as.integer(
      stringr::str_extract(as.character(rating_raw), "^-?\\d+")
    ),

    rating = if_else(
      rating %in% 1:5,
      rating,
      NA_integer_
    )
  ) %>%
  select(ID, question, vistype, rating)
likert
```
```{r likert-summary-table}
likert_table <- likert %>%
  group_by(vistype, question) %>%
  summarise(
    mean_rating = round(mean(rating, na.rm = TRUE), 2),
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(
    names_from = question,
    values_from = mean_rating
  ) %>% mutate(
    sum = rowSums(across(where(is.numeric)), na.rm = TRUE)
  ) %>%
  arrange(desc(sum))

likert_table
```
## Ordinal mixed-effects models

```{r}
likert_model_data <- likert |>
  filter(!is.na(rating)) |>
  mutate(
    ID = factor(ID),
    vistype = droplevels(factor(vistype)),
    rating_ord = ordered(rating, levels = 1:5)
  )

likert_models <- likert_model_data |>
  (\(data) split(data, data$question))() |>
  map(
    \(data_question) clmm(
      rating_ord ~ vistype + (1 | ID),
      data = data_question,
      Hess = TRUE
    )
  )

likert_estimates <- imap_dfr(
  likert_models,
  function(model, question_name) {

    emmeans(
      model,
      ~ vistype,
      mode = "mean.class"
    ) |>
      confint() |>
      as.data.frame() |>
      transmute(
        question = question_name,
        vistype,
        estimated_rating = mean.class,
        lower_ci = asymp.LCL,
        upper_ci = asymp.UCL
      )
  }
)

likert_estimates
```

## Model-estimated ratings

```{r}
#| fig-width: 14
#| fig-height: 10
question_columns <- unname(question_labels)

make_likert_estimate_plot <- function(question_name, show_x = TRUE) {

  plot_data <- likert_estimates |>
    filter(question == question_name) |>
    arrange(estimated_rating) |>
    mutate(
      colour_group = case_when(
        vistype %in% c("adjacent", "triple_CI") ~ "red",
        vistype == "basic" ~ "blue",
        TRUE ~ "other"
      ),
    
      display_label = unname(
        vistype_labels[as.character(vistype)]
      ),
    
      vistype = factor(
        vistype,
        levels = unique(vistype)
      )
    )

  p <- ggplot(
    plot_data,
      aes(
    x = estimated_rating,
    y = vistype,
    colour = colour_group
  )
  ) +
    geom_errorbar(
      aes(
        xmin = lower_ci,
        xmax = upper_ci
      ),
      orientation = "y",
      width = 0.16,
      linewidth = 0.7
    ) +
    geom_point(size = 3.5) +

    # Estimated values
    geom_text(
      aes(label = sprintf("%.2f", estimated_rating)),
      nudge_y = 0.32,
      vjust = 0.5,
      size = 5
    ) +

    # Custom y-axis labels
    geom_text(
      data = plot_data,
      aes(
        x = 0.78,
        y = vistype,
        label = display_label,
        colour = colour_group
      ),
      inherit.aes = FALSE,
      hjust = 1,
      size = 7,
      lineheight = 0.9
    ) +

    scale_colour_manual(
  values = c(
    red = "#D1495B",
    blue = "#3B78B4",
    other = "#222222"
  ),
  guide = "none"
) +

    scale_x_continuous(
      breaks = 1:5,
      labels = function(x) sprintf("%.0f", x),
      expand = expansion(mult = c(0.02, 0.02))
    ) +

    # Hide ordinary y-axis labels
    scale_y_discrete(
      labels = NULL,
      expand = expansion(add = 0.5)
    ) +

    # Show custom labels placed slightly outside the panel
    coord_cartesian(
      xlim = c(1, 5),
      clip = "off"
    ) +

    labs(
      title = question_name,
      x = "Model-estimated\nexpected rating",
      y = NULL
    ) +

    theme_classic(base_family = "Lato") +

    theme(
      plot.title = element_text(
        hjust = 0.5,
        face = "bold",
        size = 16
      ),
      axis.text.x = element_text(size = 16),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),

      # Space for custom labels
       plot.margin = margin(
    t = 5.5,
    r = 5.5,
        b = 14,
    l = 200
  )
    )

  if (!show_x) {
    p <- p +
      theme(
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()
      )
  }

  p
}

p1 <- make_likert_estimate_plot(
  question_columns[1],
  show_x = FALSE
)

p2 <- make_likert_estimate_plot(
  question_columns[2],
  show_x = FALSE
)

p3 <- make_likert_estimate_plot(
  question_columns[3],
  show_x = TRUE
)

p4 <- make_likert_estimate_plot(
  question_columns[4],
  show_x = TRUE
)

likert_estimate_figure <-
  (p1 + p2) /
  (p3 + p4) +
  plot_annotation(
    caption = paste0(
      "Points show model-estimated expected ratings.\n",
      "Error bars indicate 95% confidence intervals."
    ),
    theme = theme(
      plot.caption = element_text(
        hjust = 0.5,
        size = 10,
        face = "italic",
        family = "Lato",
        margin = margin(t = 8, b = 8)
      ),
      plot.margin = margin(6, 12, 12, 12)
    )
  )


likert_estimate_figure

if (isTRUE(params$export_artifacts)) {
  ggsave(
    "likert_estimate_figure.png",
    likert_estimate_figure,
    width = 14,
    height = 10,
    dpi = 90
  )
}
```


```{r}
likert_estimates |>
  mutate(
    ci_width = upper_ci - lower_ci,
    lower_half = estimated_rating - lower_ci,
    upper_half = upper_ci - estimated_rating
  ) |>
  arrange(question, vistype) |>
  select(
    question,
    vistype,
    estimated_rating,
    lower_ci,
    upper_ci,
    ci_width,
    lower_half,
    upper_half
  )
```

# Open-ended feedback
```{r}
feedback_A305 <- data %>%
  transmute(
    ID,
    feedback = stringr::str_squish(as.character(A305_01))
  ) %>%
  mutate(
    feedback = na_if(feedback, ""),
    feedback = na_if(feedback, "-9"),
    feedback = na_if(feedback, "NA")
  ) %>%
  filter(!is.na(feedback))
```

```{r}
feedback_A306 <- data %>%
  transmute(
    ID,
    feedback = stringr::str_squish(as.character(A306_01))
  ) %>%
  mutate(
    feedback = na_if(feedback, ""),
    feedback = na_if(feedback, "-9"),
    feedback = na_if(feedback, "NA")
  ) %>%
  filter(!is.na(feedback))
```

```{r}
feedback_A305 %>%
  mutate(
    feedback = stringr::str_wrap(feedback, width = 100)
  ) %>%
  split(.$ID) %>%
  purrr::iwalk(function(x, id) {
    cat(
      "\nID:", id, "\n",
      x$feedback,
      "\n",
      paste(rep("-", 100), collapse = ""),
      "\n",
      sep = ""
    )
  })
```

```{r}
feedback_A306 %>%
  mutate(
    feedback = stringr::str_wrap(feedback, width = 100)
  ) %>%
  split(.$ID) %>%
  purrr::iwalk(function(x, id) {
    cat(
      "\nID:", id, "\n",
      x$feedback,
      "\n",
      paste(rep("-", 100), collapse = ""),
      "\n",
      sep = ""
    )
  })
```

## Optional feedback export to PDF

```{r}
feedback_to_pdf <- function(df, title, filename) {

  df <- df %>%
    filter(!is.na(feedback), feedback != "") %>%
    arrange(ID)

  comments <- paste0(
    "<section class='comment'>",
    "<div class='id'>ID: ",
    htmltools::htmlEscape(as.character(df$ID)),
    "</div>",
    "<div class='text'>",
    htmltools::htmlEscape(df$feedback),
    "</div>",
    "</section>"
  )

  html <- paste0(
    "<!DOCTYPE html>
    <html>
    <head>
    <meta charset='UTF-8'>
    <style>
      @page {
        size: A4;
        margin: 18mm;
      }

      body {
        font-family: Arial, sans-serif;
        font-size: 11pt;
        line-height: 1.45;
        color: #222;
      }

      h1 {
        font-size: 18pt;
        margin-bottom: 20px;
      }

      .comment {
        padding: 10px 0 14px 0;
        border-bottom: 1px solid #cccccc;
        page-break-inside: avoid;
      }

      .id {
        font-weight: bold;
        margin-bottom: 5px;
      }

      .text {
        white-space: pre-wrap;
        overflow-wrap: anywhere;
      }
    </style>
    </head>

    <body>
      <h1>", htmltools::htmlEscape(title), "</h1>",
      paste(comments, collapse = "\n"),
    "</body>
    </html>"
  )

  html_file <- tempfile(fileext = ".html")
  writeLines(html, html_file, useBytes = TRUE)

  output_file <- normalizePath(filename, mustWork = FALSE)

  pagedown::chrome_print(
    input = html_file,
    output = output_file
  )

  message("Saved to: ", output_file)
}
```

```{r}
if (isTRUE(params$export_artifacts)) {
  feedback_to_pdf(
    feedback_A305,
    title = "Feedback - A305",
    filename = "feedback_A305.pdf"
  )

  feedback_to_pdf(
    feedback_A306,
    title = "Feedback - A306",
    filename = "feedback_A306.pdf"
  )
}
```

