---
title: "Open-ended feedback"
execute:
  warning: false
  message: false
params:
  export_artifacts: false
---

```{r shared-setup}
#| include: false
source("scripts/analysis_setup.R")

a305_content_analysis <- readxl::read_xlsx(
  project_path("data", "a305_content_analysis.xlsx"),
  sheet = "A305 content analysis",
  col_types = "text"
)
```

**Participants included in the qualitative content analysis:** `r n_distinct(a305_content_analysis$ID)`.

Participant comments are presented in two groups. Select an **Expand** button to read the responses.

## Qualitative content analysis

The analysis contains `r nrow(a305_content_analysis)` evaluative meaning units from `r n_distinct(a305_content_analysis$ID)` participants. `ID` links every unit to its original A305 response. Participant wording is retained as closely as possible; themes and intensities are interpretive codes. Use the search box or column filters to explore the table.

The September 22, 2026 update adds 16 meaning units from participants 111 and 112, coded with AI assistance using the existing codebook. The previous 224 meaning units are unchanged. The source survey contains `r nrow(data)` eligible participants, with responses through `r max(as.Date(data$STARTED))`.

For the new responses, references to bivariate plots without an identified variant are coded as `General`. Proposed redesigns are included only when they express an evaluation of the displayed visualization; suggestions alone are not assigned an intensity.

[Download the coded content analysis as an Excel workbook](../data/a305_content_analysis.xlsx)

### Theme codebook

```{r a305-theme-codebook}
#| echo: false

tribble(
  ~Theme, ~Explanation,
  "Accessibility", "Usability across color vision, devices, or other viewer needs.",
  "Aesthetics", "Visual attractiveness, cleanliness, or perceived ugliness.",
  "Information integration", "How well multiple quantities are combined without obscuring one another.",
  "Informativeness", "Amount or usefulness of information communicated.",
  "Interpretability", "Ease or difficulty of understanding what the visualization means.",
  "Learnability", "Whether interpretation becomes easier with exposure or practice.",
  "Legend/color mapping", "Clarity and intuitiveness of colors, scales, and their legends.",
  "Medium compatibility", "Suitability for print, papers, presentations, screens, or animation.",
  "Preference", "An explicit choice, ranking, liking, or dislike.",
  "Reading effort", "Time, attention, or cognitive effort required to extract information.",
  "SNR visibility", "How clearly signal-to-noise ratio or uncertainty can be identified.",
  "Space/scalability", "Figure size, visual density, or ability to accommodate multiple plots.",
  "Task suitability", "Usefulness for a particular analytical or comparison task."
) |>
  knitr::kable(
    format = "html",
    col.names = c("Theme", "Explanation"),
    table.attr = kable_table_attributes
  )
```

### Coded meaning units

```{r a305-content-analysis}
#| echo: false

a305_content_analysis |>
  select(ID, Visualization, `Meaning unit`, Theme, `Evaluation intensity`) |>
  reactable::reactable(
    searchable = TRUE,
    filterable = TRUE,
    striped = TRUE,
    highlight = TRUE,
    bordered = TRUE,
    defaultPageSize = 20,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(10, 20, 50, 100),
    columns = list(
      ID = reactable::colDef(
        name = "Participant ID",
        minWidth = 70,
        align = "center"
      ),
      Visualization = reactable::colDef(
        name = "Visualization type",
        minWidth = 145
      ),
      `Meaning unit` = reactable::colDef(minWidth = 420),
      Theme = reactable::colDef(minWidth = 160),
      `Evaluation intensity` = reactable::colDef(
        minWidth = 150,
        align = "center"
      )
    )
  )
```

## Evaluation intensity by theme and visualization type

This descriptive analysis gives each participant at most one value within each visualization–theme combination. It therefore avoids treating multiple meaning units from the same participant as independent observations. Comments coded as `General` are retained in the original dataset but excluded here because they do not identify one of the seven visualization types.

```{r a305-participant-theme-aggregation}
#| include: false

visualization_order <- c(
  "Basic",
  "Adjacent",
  "Bivariate corner",
  "Bivariate range",
  "Confidence interval",
  "Uncertainty markers",
  "Animation"
)

a305_participant_theme <- a305_content_analysis |>
  transmute(
    participant = ID,
    visualization = factor(Visualization, levels = visualization_order),
    theme = Theme,
    intensity = as.numeric(gsub("+", "", `Evaluation intensity`, fixed = TRUE))
  ) |>
  filter(!is.na(visualization), !is.na(theme), !is.na(intensity)) |>
  group_by(participant, visualization, theme) |>
  summarise(
    n_units = n(),
    median_intensity = median(intensity),
    .groups = "drop"
  )

a305_intensity_summary <- a305_participant_theme |>
  rename(participant_median_intensity = median_intensity) |>
  group_by(theme, visualization) |>
  summarise(
    n_participants = n_distinct(participant),
    median_intensity = median(participant_median_intensity),
    mean_intensity = mean(participant_median_intensity),
    .groups = "drop"
  ) |>
  arrange(theme, visualization)
```

```{r a305-intensity-theme-figure}
#| fig-width: 15
#| fig-height: 19
#| out-width: 100%
#| fig-alt: "Faceted boxplots of participant-level median evaluation intensity by visualization type and qualitative theme, with individual participant points."

intensity_axis_labels <- c(
  `-2` = "−2",
  `-1` = "−1",
  `0` = "0",
  `1` = "+1",
  `2` = "+2"
)

visualization_axis_labels <- c(
  Basic = "<span style='color:#377EB8'>Basic</span>",
  Adjacent = "Adjacent",
  `Bivariate corner` = "Bivariate<br>corner",
  `Bivariate range` = "Bivariate<br>range",
  `Confidence interval` = "Confidence<br>interval",
  `Uncertainty markers` = "Uncertainty<br>markers",
  Animation = "Animation"
)

theme_panel_counts <- a305_participant_theme |>
  count(theme, name = "n_entries") |>
  arrange(desc(n_entries), theme)

theme_panel_order <- theme_panel_counts$theme

theme_panel_labels <- setNames(
  paste0(
    theme_panel_counts$theme,
    " (N = ",
    theme_panel_counts$n_entries,
    ")"
  ),
  theme_panel_counts$theme
)

a305_intensity_plot_data <- a305_participant_theme |>
  mutate(theme = factor(theme, levels = theme_panel_order))

a305_intensity_by_theme_figure <- ggplot(
  a305_intensity_plot_data,
  aes(
    x = visualization,
    y = median_intensity,
    colour = visualization == "Basic",
    fill = visualization == "Basic"
  )
) +
  geom_boxplot(
    width = 0.62,
    outlier.shape = NA,
    linewidth = 0.6
  ) +
  geom_point(
    position = position_jitter(
      width = 0.16,
      height = 0.06,
      seed = 2025
    ),
    size = 2.4,
    alpha = 0.8
  ) +
  facet_wrap(
    vars(theme),
    scales = "free_x",
    ncol = 2,
    labeller = as_labeller(theme_panel_labels)
  ) +
  scale_x_discrete(labels = visualization_axis_labels, drop = TRUE) +
  scale_colour_manual(
    values = c(`FALSE` = "#333333", `TRUE` = "#377EB8"),
    guide = "none"
  ) +
  scale_fill_manual(
    values = c(`FALSE` = "#D9D9D9", `TRUE` = "#B9D8F2"),
    guide = "none"
  ) +
  scale_y_continuous(
    limits = c(-2.15, 2.15),
    breaks = -2:2,
    labels = intensity_axis_labels,
    minor_breaks = NULL
  ) +
  labs(
    x = NULL,
    y = "Participant-level median evaluation intensity",
    caption = paste(
      "Points represent participants, not individual meaning units.",
      "Panel N is the number of plotted participant–visualization entries.",
      "−2 = strongly negative; +2 = strongly positive."
    )
  ) +
  theme_classic(base_family = "Lato", base_size = 18) +
  theme(
    strip.text = element_text(face = "bold", size = 40),
    axis.text.x = ggtext::element_markdown(
      size = 24,
      lineheight = 0.85,
      margin = margin(t = 6)
    ),
    axis.text.y = element_text(size = 16),
    axis.title.y = element_text(size = 40),
    axis.ticks.x = element_blank(),
    panel.spacing = unit(1.2, "lines"),
    plot.caption = element_text(hjust = 0.5, face = "italic", size = 14)
  )

a305_intensity_by_theme_figure

ggsave(
  project_path("figures", "a305_intensity_by_theme_figure.png"),
  a305_intensity_by_theme_figure,
  width = 15,
  height = 19,
  units = "in",
  dpi = 180,
  device = ragg::agg_png,
  bg = "white"
)
```

### Participant-based summary

Sample sizes count unique participants. Medians and means are calculated from the participant-level medians shown in the figure, not directly from meaning units.

```{r a305-intensity-summary-table}
a305_intensity_summary |>
  mutate(
    visualization = as.character(visualization),
    median_intensity = round(median_intensity, 2),
    mean_intensity = round(mean_intensity, 2)
  ) |>
  select(
    theme,
    visualization,
    n_participants,
    median_intensity,
    mean_intensity
  ) |>
  reactable::reactable(
    searchable = TRUE,
    filterable = TRUE,
    striped = TRUE,
    highlight = TRUE,
    bordered = TRUE,
    defaultPageSize = 20,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(10, 20, 50, 100),
    columns = list(
      theme = reactable::colDef(name = "Theme", minWidth = 170),
      visualization = reactable::colDef(
        name = "Visualization type",
        minWidth = 160
      ),
      n_participants = reactable::colDef(
        name = "Participants",
        align = "center"
      ),
      median_intensity = reactable::colDef(
        name = "Median intensity",
        align = "center",
        format = reactable::colFormat(digits = 2)
      ),
      mean_intensity = reactable::colDef(
        name = "Mean intensity",
        align = "center",
        format = reactable::colFormat(digits = 2)
      )
    )
  )
```

## Original responses

```{r}
feedback_A305 <- data %>%
  transmute(
    ID,
    feedback = stringr::str_squish(as.character(A305_01))
  ) %>%
  mutate(
    feedback = na_if(feedback, ""),
    feedback = na_if(feedback, "-9"),
    feedback = na_if(feedback, "NA")
  ) %>%
  filter(!is.na(feedback))
```

```{r}
feedback_A306 <- data %>%
  transmute(
    ID,
    feedback = stringr::str_squish(as.character(A306_01))
  ) %>%
  mutate(
    feedback = na_if(feedback, ""),
    feedback = na_if(feedback, "-9"),
    feedback = na_if(feedback, "NA")
  ) %>%
  filter(!is.na(feedback))
```

Comments specifically about the visualizations used in the study.

::: {.callout-note collapse="true"}
## Visualization feedback — Expand

```{r}
feedback_A305 %>%
  mutate(
    feedback = stringr::str_wrap(feedback, width = 100)
  ) %>%
  split(.$ID) %>%
  purrr::iwalk(function(x, id) {
    cat(
      "\nID:", id, "\n",
      x$feedback,
      "\n",
      paste(rep("-", 100), collapse = ""),
      "\n",
      sep = ""
    )
  })
```

:::

Broader comments about the study, tasks, or participants' experience.

::: {.callout-note collapse="true"}
## General feedback — Expand

```{r}
feedback_A306 %>%
  mutate(
    feedback = stringr::str_wrap(feedback, width = 100)
  ) %>%
  split(.$ID) %>%
  purrr::iwalk(function(x, id) {
    cat(
      "\nID:", id, "\n",
      x$feedback,
      "\n",
      paste(rep("-", 100), collapse = ""),
      "\n",
      sep = ""
    )
  })
```

:::

## Optional feedback export to PDF

```{r}
feedback_to_pdf <- function(df, title, filename) {

  df <- df %>%
    filter(!is.na(feedback), feedback != "") %>%
    arrange(ID)

  comments <- paste0(
    "<section class='comment'>",
    "<div class='id'>ID: ",
    htmltools::htmlEscape(as.character(df$ID)),
    "</div>",
    "<div class='text'>",
    htmltools::htmlEscape(df$feedback),
    "</div>",
    "</section>"
  )

  html <- paste0(
    "<!DOCTYPE html>
    <html>
    <head>
    <meta charset='UTF-8'>
    <style>
      @page {
        size: A4;
        margin: 18mm;
      }

      body {
        font-family: Arial, sans-serif;
        font-size: 11pt;
        line-height: 1.45;
        color: #222;
      }

      h1 {
        font-size: 18pt;
        margin-bottom: 20px;
      }

      .comment {
        padding: 10px 0 14px 0;
        border-bottom: 1px solid #cccccc;
        page-break-inside: avoid;
      }

      .id {
        font-weight: bold;
        margin-bottom: 5px;
      }

      .text {
        white-space: pre-wrap;
        overflow-wrap: anywhere;
      }
    </style>
    </head>

    <body>
      <h1>", htmltools::htmlEscape(title), "</h1>",
      paste(comments, collapse = "\n"),
    "</body>
    </html>"
  )

  html_file <- tempfile(fileext = ".html")
  writeLines(html, html_file, useBytes = TRUE)

  output_file <- normalizePath(filename, mustWork = FALSE)

  pagedown::chrome_print(
    input = html_file,
    output = output_file
  )

  message("Saved to: ", output_file)
}
```

```{r}
if (isTRUE(params$export_artifacts)) {
  feedback_to_pdf(
    feedback_A305,
    title = "Feedback - A305",
    filename = "feedback_A305.pdf"
  )

  feedback_to_pdf(
    feedback_A306,
    title = "Feedback - A306",
    filename = "feedback_A306.pdf"
  )
}
```
