---
title: "Sample"
execute:
  warning: false
  message: false
params:
  export_artifacts: false
---

```{r shared-setup}
#| include: false
source("scripts/analysis_setup.R")
```

## Included sample: N = `r nrow(data)`

The analysis includes `r nrow(data)` respondents after applying the study's
date and completion criteria.

```{r sample-summary}
tibble(
  respondents = nrow(data),
  mean_completion_minutes = mean(as.numeric(data$TIME_SUM), na.rm = TRUE) / 60,
  median_completion_minutes = median(as.numeric(data$TIME_SUM), na.rm = TRUE) / 60
)
```

# Participant characteristics

## Self-assessed visualization experience
```{r}
n_total <- data %>%
  filter(!is.na(A001)) %>%
  nrow()

data$A001 %>% table() %>%  data.frame() %>% dplyr::rename(level = !!names(.)[1]) %>%
    mutate(percent_score = round(Freq / sum(Freq) * 100)) %>% 
    mutate(level = factor(level, levels = rev(c("Beginner", "Intermediate", "Expert")))) %>% 
    ggplot(data = ., aes(y = level, x = percent_score)) +  
    geom_col(stat="identity",  width = 0.8, fill ="#6BAED6") + labs(x = "Percent of respondents", y = "", title="Self-assessed level") +
    geom_text(aes(label = paste0(percent_score, "%")), hjust = -0.2)  +  theme_classic(base_family = "Lato", base_size = 20) + 
 scale_x_continuous(
    name = "Percent of respondents",
    labels = \(x) paste0(x, "%"),
    breaks = seq(0, 50, 10),

    sec.axis = sec_axis(
      transform = ~ . * n_total / 100,
      name = "Number of respondents"
    ),

    expand = expansion(mult = c(0, 0.12))
  ) + theme(
    plot.title = element_text(
      size = 22,
      face = "bold", hjust = 0.5,
      margin = margin(b = 14)
    ),
    axis.title.x = element_text(
      size = 20, hjust = 0.5,
      margin = margin(t = 10),
    ),
    axis.text = element_text(size = 20),
    axis.ticks.y = element_blank(),
    legend.position = "none"
  )  + coord_cartesian(clip = "off")
```

## Country of work
```{r}
country <- data %>%
  transmute(
    country_work = case_when(
      A004 == "-2" & !is.na(A004s) & A004s != "" ~ A004s,
      TRUE ~ A004
    )
  ) %>% mutate(
    country_work = trimws(country_work),

    country_work = case_when(
      country_work %in% c("NL", "The Netherlands") ~ "Netherlands",
      country_work %in% c("United States", "United States of America") ~ "USA",
      country_work %in% c(
        "Prefer not to say",
        "Professor",
        "Senior Researcher"
      ) ~ NA_character_,
      TRUE ~ country_work
    )) %>%
  filter(!is.na(country_work), country_work != "") %>%
  mutate(
    country_work = fct_lump_min(
      factor(country_work),
      min = 2,
      other_level = "Other"
    )
  ) %>% count(country_work, name = "Freq") %>%
  mutate(
    percent_score = round(Freq / sum(Freq) * 100),
    country_work = fct_reorder(
      country_work,
      percent_score
    )
  )
country

```

### Diagnostic: responses grouped as “Other”

```{r country-other-diagnostic}
country_other <- data %>%
  transmute(
    country_original = case_when(
      A004 == "-2" & !is.na(A004s) & A004s != "" ~ A004s,
      TRUE ~ A004
    )
  ) %>%
  mutate(
    country_original = trimws(country_original),
    country_original = case_when(
      country_original %in% c("NL", "The Netherlands") ~ "Netherlands",
      country_original %in% c("United States", "United States of America") ~ "USA",
      country_original %in% c(
        "Prefer not to say",
        "Professor",
        "Senior Researcher"
      ) ~ NA_character_,
      TRUE ~ country_original
    )
  ) %>%
  filter(!is.na(country_original), country_original != "") %>%
  count(country_original, name = "respondents", sort = TRUE) %>%
  filter(respondents < 2)

country_other
```

```{r}

country %>%
  ggplot(aes(percent_score, reorder(country_work, percent_score))) +
  geom_col(width = 0.7, fill = "#6BAED6") +
  geom_text(aes(label = paste0(round(percent_score), "%")),
            hjust = -0.15, size = 6) +
  scale_x_continuous(
    "Percent of respondents",
    labels = \(x) paste0(x, "%"),
    sec.axis = sec_axis(~ . * n_total / 100, name = "Number of respondents"),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(
    title = "Country of work",
    y = NULL,
   # caption = paste("N of respondents =", n_total)
  ) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 22),
    axis.text = element_text(size = 20),
    axis.ticks.y = element_blank(),
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0)
  )
```

## Research field
```{r}
field <- data %>%
  mutate(field = stringr::word(A005, 1)) %>%
  count(field, name = "Freq") %>%
  mutate(percent_score = Freq / sum(Freq) * 100)
field
```
```{r}
field %>%
  ggplot(aes(percent_score, reorder(field, percent_score))) +
  geom_col(width = 0.7, fill = "#6BAED6") +
  geom_text(aes(label = paste0(round(percent_score), "%")),
            hjust = -0.15, size = 6) +
  scale_x_continuous(
    "Percent of respondents",
    labels = \(x) paste0(x, "%"),
    sec.axis = sec_axis(~ . * n_total / 100,
                        name = "Number of respondents"),
    expand = expansion(mult = c(0, 0.15))
  ) +
  labs(title = "Research field", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 22),
    axis.text = element_text(size = 20),
    axis.ticks.y = element_blank()
  )
```
## Software used

```{r}
software_names <- legend %>%
  select(matches("^A006_\\d{2}$")) %>%
  pivot_longer(everything(), names_to = "item", values_to = "software")  %>%
  mutate(
    software = stringr::word(software, 2, -1),
    software = if_else(item == "A006_27", "Other", software)
  )


software <- data %>%
  select(matches("^A006_\\d{2}$")) %>%
  pivot_longer(everything(), names_to = "item", values_to = "response") %>%
  filter(response == "Checked") %>%
  count(item, name = "Freq") %>%
  right_join(software_names, by = "item") %>%
  mutate(
    Freq = coalesce(Freq, 0L),
    percent_score = Freq / n_total * 100
  )
software
```
```{r}
software %>% filter(Freq != 0) %>% 
  ggplot(aes(percent_score, reorder(software, percent_score))) +
  geom_col(width = 0.7, fill = "#6BAED6") +
  geom_text(aes(label = paste0(round(percent_score), "%")),
            hjust = -0.15, size = 6) +
  scale_x_continuous(
    "Percent of respondents",
    labels = \(x) paste0(x, "%"),
    sec.axis = sec_axis(~ . * n_total / 100,
                        name = "Number of respondents"),
    expand = expansion(mult = c(0, 0.18))
  ) +
  labs(title = "Software used", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 22),
    axis.text = element_text(size = 20),
    axis.ticks.y = element_blank()
  )
```

## Combined participant-characteristics figure

```{r participant-characteristics-figure}
#| fig-width: 15
#| fig-height: 13

format_category_labels <- function(labels) {
  lapply(
    labels,
    \(label) {
      if (identical(label, "Other")) {
        bquote(italic(.(label)))
      } else {
        label
      }
    }
  )
}

make_characteristic_plot <- function(
  plot_data,
  label_column,
  title,
  bar_width = 0.7,
  label_size = 20
) {
  plot_data <- plot_data %>%
    filter(!is.na(.data[[label_column]])) %>%
    mutate(
      display_label = forcats::fct_reorder(
        factor(.data[[label_column]]),
        percent_score
      )
    )

  ggplot(plot_data, aes(x = percent_score, y = display_label)) +
    geom_col(width = bar_width, fill = "#6BAED6") +
    geom_text(
      aes(label = paste0(round(percent_score), "%")),
      hjust = -0.15,
      size = 8
    ) +
    scale_x_continuous(
      limits = c(0, 100),
      breaks = seq(0, 100, 25),
      labels = \(x) paste0(x, "%"),
      expand = expansion(mult = c(0, 0.02))
    ) +
    scale_y_discrete(labels = format_category_labels) +
    labs(title = title, x = "Percent of respondents", y = NULL) +
    coord_cartesian(clip = "off") +
    theme_classic(base_family = "Lato", base_size = 20) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 22),
      axis.title.x = element_text(size = 20),
      axis.text.x = element_text(size = 20),
      axis.text.y = element_text(size = label_size),
      axis.ticks.y = element_blank(),
      plot.margin = margin(8, 20, 8, 8)
    )
}

experience_combined <- data %>%
  filter(!is.na(A001), A001 != "") %>%
  count(level = A001, name = "Freq") %>%
  mutate(percent_score = Freq / sum(Freq) * 100)

country_combined <- country %>%
  transmute(
    country_work = as.character(country_work),
    Freq,
    percent_score
  )

field_combined <- field %>%
  filter(!is.na(field), field != "")

software_combined <- software %>%
  filter(Freq > 0, !is.na(software), software != "")

p_experience <- make_characteristic_plot(
  experience_combined,
  "level",
  "Self-assessed level"
)
p_country <- make_characteristic_plot(
  country_combined,
  "country_work",
  "Country of work"
)
p_field <- make_characteristic_plot(
  field_combined,
  "field",
  "Research field"
)
p_software <- make_characteristic_plot(
  software_combined,
  "software",
  "Software used"
)

participant_characteristics_figure <-
  (p_experience + p_field) /
  (p_country + p_software) +
  plot_layout(heights = c(0.65, 1.35)) +
  plot_annotation(
    tag_levels = "A",
    caption = paste0(
      "Bars show percentages of respondents with valid responses. ",
      "Multiple software selections were permitted."
    ),
    theme = theme(
      plot.tag = element_text(face = "bold", size = 18),
      plot.caption = element_text(
        hjust = 0.5,
        face = "italic",
        family = "Lato",
        size = 18
      )
    )
  )

participant_characteristics_figure

if (isTRUE(params$export_artifacts)) {
  ggsave(
    project_path("figures", "participant_characteristics_figure.png"),
    participant_characteristics_figure,
    width = 15,
    height = 13,
    dpi = 140
  )
}
```

### Reduced figure excluding categories at or below 3%

```{r participant-characteristics-figure-over-3-percent}
#| fig-width: 15
#| fig-height: 9

country_over_3 <- country_combined %>%
  mutate(percent_score = Freq / sum(Freq) * 100) %>%
  filter(round(percent_score) > 3)

software_over_3 <- software_combined %>%
  filter(round(percent_score) > 3)

p_country_over_3 <- make_characteristic_plot(
  country_over_3,
  "country_work",
  "Country of work",
  label_size = 24
)

p_software_over_3 <- make_characteristic_plot(
  software_over_3,
  "software",
  "Software used"
)

participant_characteristics_over_3_figure <-
  (p_experience + p_field) /
  (p_country_over_3 + p_software_over_3) +
  plot_layout(heights = c(0.65, 1.35)) +
  plot_annotation(
    tag_levels = "A",
    caption = paste0(
      "Categories ≥ 3% are combined into 'Other' in panels C and D. ",
      "Multiple software selections were permitted."
    ),
    theme = theme(
      plot.tag = element_text(face = "bold", size = 18),
      plot.caption = element_text(
        hjust = 0.5,
        face = "italic",
        family = "Lato",
        size = 18
      )
    )
  )

participant_characteristics_over_3_figure

if (isTRUE(params$export_artifacts)) {
  ggsave(
    project_path("figures", "participant_characteristics_over_3_figure.png"),
    participant_characteristics_over_3_figure,
    width = 15,
    height = 13,
    dpi = 140
  )
}

```


