---
title: "Task 1"
execute:
  warning: false
  message: false
params:
  export_artifacts: false
---

```{r shared-setup}
#| include: false
source("scripts/analysis_setup.R")
```


## Trial reconstruction and presentation order
```{r}
task1_order <- data %>%
  dplyr::select(ID, A125_02) %>%
  tidyr::separate_rows(A125_02, sep = "\n") %>%
  dplyr::filter(A125_02 != "") %>%
  dplyr::transmute(
    ID,
    position = as.integer(get_field(A125_02, "position")),
    trial    = as.integer(get_field(A125_02, "trial")),
    question = get_field(A125_02, "question"),
    stimulus = get_field(A125_02, "stimulus"),
    expected = as.integer(get_field(A125_02, "expected")),
    answered = as.integer(get_field(A125_02, "answered")),
    correct  = as.integer(get_field(A125_02, "correct"))
  ) %>%
  dplyr::arrange(ID, position)

task1_design <- data %>%
  dplyr::select(ID, A124_01) %>%
  tidyr::separate_rows(A124_01, sep = "\n") %>%
  dplyr::filter(A124_01 != "") %>%
  dplyr::transmute(
    ID,
    trial       = as.integer(get_field(A124_01, "trial")),
    question    = get_field(A124_01, "question"),
    vistype     = get_field(A124_01, "vistype"),
    target      = get_field(A124_01, "target"),
    correct     = get_field(A124_01, "correct"),
    wrong       = get_field(A124_01, "wrong"),
    pair        = get_field(A124_01, "pair"),
    solution    = as.integer(get_field(A124_01, "solution")),
    opt1        = get_field(A124_01, "opt1"),
    opt2        = get_field(A124_01, "opt2")
  )

task1 <- task1_order %>%
  left_join(
    task1_design %>%
      select(ID, trial, question, vistype),
    by = c("ID", "trial", "question")
  )
```

## Accuracy analysis
```{r}
task1_model <- task1 %>%
  transmute(
    ID = factor(ID),
    trial = factor(trial),
    vistype = relevel(factor(vistype), ref = "basic"),
    position = as.integer(position),
    correct = as.integer(correct)
  ) %>%
  filter(!is.na(correct), !is.na(position), !is.na(vistype)) %>%
  mutate(position_c = position - mean(position))
```

```{r}
vistype_table <- task1_model |>
  group_by(vistype) |>
  summarise(
    n = n(),
    n_correct = sum(correct == 1, na.rm = TRUE),
    n_incorrect = sum(correct == 0, na.rm = TRUE),
    accuracy = mean(correct, na.rm = TRUE),
    sd_accuracy = sd(correct, na.rm = TRUE),
    .groups = "drop"
  )
vistype_table |>
  mutate(
    vistype = table_vistype(vistype),
    accuracy = scales::percent(accuracy, accuracy = 0.1),
    sd_accuracy = scales::percent(sd_accuracy, accuracy = 0.1)
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Visualization type",
      "Responses",
      "Correct",
      "Incorrect",
      "Mean accuracy",
      "SD"
    ),
    caption = "Observed comparison-task accuracy by visualization type",
    table.attr = kable_table_attributes
  )
```
```{r}
vistype_table |>
  ggplot(aes(accuracy, reorder(vistype, accuracy))) +
  geom_col(
    width = 0.7,
    fill = "#6BAED6"
  ) +
  geom_errorbar(
    aes(
      xmin = pmax(accuracy - sd_accuracy, 0),
      xmax = pmin(accuracy + sd_accuracy, 1)
    ),
    orientation = "y",
    width = 0.14,
    linewidth = 0.7,
    position = position_nudge(y = 0.17)
  ) +
  geom_text(
    aes(label = sprintf("%.2f", accuracy)),
    hjust = -0.15,
    nudge_y = -0.20,
    size = 6
  ) +
  scale_x_continuous(
    "Mean accuracy",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    labels = scales::label_percent(),
    expand = expansion(mult = c(0, 0.10))
  ) +
  scale_y_discrete(
    labels = vistype_labels
  ) +
  labs(
    title = "Comparison task accuracy",
    y = NULL,
    caption = "Bars show mean accuracy; error bars indicate ±1 SD."
  ) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 22
    ),
    axis.text = element_text(size = 20),
    axis.ticks.y = element_blank(),
    plot.caption = element_text(
      hjust = 0.5,
      face = "italic"
    )
  )
```

```{r}

# Logistic mixed-effects model
task1_fit <- glmer(
  correct ~ vistype + position_c + (1 | ID),
  data = task1_model,
  family = binomial
)

# Estimated probabilities and 95% confidence intervals
accuracy_estimates <- emmeans(
  task1_fit,
  ~ vistype,
  type = "response"
) |>
  confint() |>
  as.data.frame() |>
  transmute(
    vistype,
    accuracy = prob,
    lower_ci = asymp.LCL,
    upper_ci = asymp.UCL
  ) |>
  arrange(desc(accuracy)) |>
  mutate(
    vistype = factor(vistype, levels = rev(vistype))
  )

accuracy_estimates |>
  mutate(
    vistype = table_vistype(vistype),
    across(
      c(accuracy, lower_ci, upper_ci),
      \(x) scales::percent(x, accuracy = 0.1)
    )
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Visualization type",
      "Estimated accuracy",
      "Lower 95% CI",
      "Upper 95% CI"
    ),
    caption = "Model-estimated comparison-task accuracy",
    table.attr = kable_table_attributes
  )
```
```{r}
# Model-estimated probability of a correct response for each visualization type.
comparison_axis_colours <- ifelse(
  levels(accuracy_estimates$vistype) == "basic",
  "#377EB8",
  "#222222"
)

comparison_accuracy_figure <- ggplot(
  accuracy_estimates,
  aes(
    x = accuracy,
    y = vistype,
    colour = vistype == "basic"
  )
) +
  geom_errorbar(
    aes(
      xmin = lower_ci,
      xmax = upper_ci
    ),
    orientation = "y",
    width = 0.14,
    linewidth = 0.55
  ) +
  geom_point(size = 3) +
  geom_text(
    aes(label = scales::percent(accuracy, accuracy = 1)),
    nudge_y = 0.36,
    vjust = 0.5,
    size = 5.5
  ) +
  scale_y_discrete(
    expand = expansion(add = 0.5),
    labels = function(x) {
      publication_labels <- c(
        basic = "Basic topoplot",
        adjacent = "Adjacent plots",
        bivariate_corner = "Bivariate corners",
        bivariate_range = "Bivariate range",
        triple_CI = "Confidence intervals",
        uncert_markers = "Uncertainty markers",
        animation = "Animation"
      )
      unname(publication_labels[x])
    }
  ) +
  scale_colour_manual(
    values = c(`FALSE` = "#222222", `TRUE` = "#377EB8"),
    guide = "none"
  ) +
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(0.65, 0.85),
    breaks = seq(0.65, 0.85, 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    title = "Comparison task accuracy",
    subtitle = "Model-estimated accuracy with 95% confidence intervals",
    x = "Model-estimated accuracy",
    y = NULL
  ) +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 18),
    plot.title.position = "plot",
    axis.text.y = element_text(
      size = 24,
      colour = comparison_axis_colours,
      margin = margin(r = 12)
    ),
    axis.text.x = element_text(size = 20),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

comparison_accuracy_figure

ggsave(
  project_path("figures", "comparison_accuracy_figure.png"),
  comparison_accuracy_figure,
  width = 9,
  height = 5.5,
  units = "in",
  dpi = 180,
  device = ragg::agg_png,
  bg = "white"
)
```

### Accuracy differences relative to Basic topoplot

This plot shows response-scale differences from Basic topoplot. Confidence intervals crossing zero indicate that the corresponding visualization is not clearly different from Basic at the 95% confidence level.

```{r task1-basic-contrast-figure}
task1_response_emmeans <- emmeans(
  task1_fit,
  ~ vistype,
  type = "response",
  regrid = "response"
)

task1_basic_reference <- which(
  task1_response_emmeans@levels$vistype == "basic"
)

task1_basic_contrasts_raw <- contrast(
  task1_response_emmeans,
  method = "trt.vs.ctrl",
  ref = task1_basic_reference
) |>
  confint() |>
  as.data.frame()

task1_lower_ci_column <- grep(
  "lower.CL$|asymp.LCL$",
  names(task1_basic_contrasts_raw),
  value = TRUE
)[1]
task1_upper_ci_column <- grep(
  "upper.CL$|asymp.UCL$",
  names(task1_basic_contrasts_raw),
  value = TRUE
)[1]

task1_basic_contrasts <- task1_basic_contrasts_raw |>
  transmute(
    vistype = stringr::str_remove(contrast, " - basic$"),
    difference = estimate,
    lower_ci = .data[[task1_lower_ci_column]],
    upper_ci = .data[[task1_upper_ci_column]]
  ) |>
  mutate(
    vistype = factor(vistype, levels = levels(accuracy_estimates$vistype))
  )

task1_contrast_limit <- max(
  abs(c(task1_basic_contrasts$lower_ci, task1_basic_contrasts$upper_ci)),
  na.rm = TRUE
)
task1_contrast_limit <- max(
  0.05,
  ceiling(task1_contrast_limit / 0.05) * 0.05
)

task1_contrast_labels <- c(
  adjacent = "Adjacent plots",
  bivariate_corner = "Bivariate corners",
  bivariate_range = "Bivariate range",
  triple_CI = "Confidence intervals",
  uncert_markers = "Uncertainty markers",
  animation = "Animation"
)

comparison_accuracy_contrast_figure <- ggplot(
  task1_basic_contrasts,
  aes(x = difference, y = vistype)
) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    linewidth = 0.6,
    colour = "#377EB8"
  ) +
  geom_errorbar(
    aes(xmin = lower_ci, xmax = upper_ci),
    orientation = "y",
    width = 0.14,
    linewidth = 0.55,
    colour = "#333333"
  ) +
  geom_point(size = 3, colour = "#333333") +
  geom_text(
    aes(label = ifelse(
      abs(difference) < 0.005,
      "0 pp",
      sprintf("%+.0f pp", 100 * difference)
    )),
    nudge_y = 0.32,
    vjust = 0.5,
    size = 5.5,
    colour = "#333333"
  ) +
  scale_y_discrete(
    labels = function(x) unname(task1_contrast_labels[x]),
    expand = expansion(add = 0.5)
  ) +
  scale_x_continuous(
    limits = c(-task1_contrast_limit, task1_contrast_limit),
    labels = function(x) ifelse(
      abs(x) < 0.005,
      "0 pp",
      sprintf("%+.0f pp", 100 * x)
    ),
    breaks = scales::breaks_pretty(n = 5),
    expand = expansion(mult = 0.04)
  ) +
  labs(
    title = "Comparison task accuracy differences",
    subtitle = "Response-scale contrasts with 95% confidence intervals",
    x = NULL,
    y = NULL
  ) +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 22),
    plot.subtitle = element_text(hjust = 0.5, size = 18),
    plot.title.position = "plot",
    axis.text.y = element_text(size = 24, margin = margin(r = 12)),
    axis.text.x = element_text(size = 20),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

comparison_accuracy_contrast_caption <- ggplot() +
  annotate(
    "text",
    x = 0.497,
    y = 0.5,
    label = "Difference in estimated accuracy vs",
    hjust = 1,
    size = 20 / ggplot2::.pt,
    family = "Lato",
    colour = "#222222"
  ) +
  annotate(
    "text",
    x = 0.503,
    y = 0.5,
    label = "Basic topoplot",
    hjust = 0,
    size = 20 / ggplot2::.pt,
    family = "Lato",
    colour = "#377EB8"
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
  theme_void()

comparison_accuracy_contrast_figure <-
  comparison_accuracy_contrast_figure /
  comparison_accuracy_contrast_caption +
  plot_layout(heights = c(1, 0.10))

comparison_accuracy_contrast_figure

ggsave(
  project_path("figures", "comparison_accuracy_contrast_figure.png"),
  comparison_accuracy_contrast_figure,
  width = 9,
  height = 5.25,
  units = "in",
  dpi = 180,
  device = ragg::agg_png,
  bg = "white"
)
```

```{r}
model_order <- glmmTMB(
  correct ~ vistype + position_c +
    (1 | ID),
  family = binomial,
  data = task1_model
)

summary(model_order)$coefficients$cond |>
  as.data.frame() |>
  tibble::rownames_to_column("term") |>
  transmute(
    term = dplyr::recode(
      term,
      `(Intercept)` = "Basic topoplot (intercept)",
      vistypeadjacent = "Adjacent plots vs Basic topoplot",
      vistypeanimation = "Animation vs Basic topoplot",
      vistypebivariate_corner = "Bivariate corner vs Basic topoplot",
      vistypebivariate_range = "Bivariate range vs Basic topoplot",
      vistypetriple_CI = "Confidence intervals vs Basic topoplot",
      vistypeuncert_markers = "Uncertainty markers vs Basic topoplot",
      position_c = "Presentation position (centered)",
      .default = term
    ),
    estimate = Estimate,
    standard_error = .data[["Std. Error"]],
    z_value = .data[["z value"]],
    p_value = format.pval(
      .data[["Pr(>|z|)"]],
      digits = 3,
      eps = 0.001
    )
  ) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Term",
      "Estimate (log odds)",
      "Standard error",
      "z value",
      "p value"
    ),
    caption = "Fixed effects for the comparison-task accuracy model",
    table.attr = kable_table_attributes
  )
```
```{r}
model_no_vistype <- glmmTMB(
  correct ~ position_c + (1 | ID),
  family = binomial,
  data = task1_model
)

model_vistype <- glmmTMB(
  correct ~ vistype + position_c + (1 | ID),
  family = binomial,
  data = task1_model
)

anova(model_no_vistype, model_vistype) |>
  as.data.frame() |>
  tibble::rownames_to_column("model") |>
  mutate(
    model = dplyr::recode(
      model,
      model_no_vistype = "Presentation position only",
      model_vistype = "Presentation position and visualization type",
      .default = model
    ),
    `Pr(>Chisq)` = format.pval(
      .data[["Pr(>Chisq)"]],
      digits = 3,
      eps = 0.001
    )
  ) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Model",
      "Parameters",
      "AIC",
      "BIC",
      "Log likelihood",
      "Deviance",
      "Likelihood-ratio chi-square",
      "df difference",
      "p value"
    ),
    caption = "Likelihood-ratio test of visualization type for comparison-task accuracy",
    table.attr = kable_table_attributes
  )
```
## Completion-time analysis

```{r}
task1_time <- data %>%
  select(ID, TIME005:TIME025) %>%
  pivot_longer(-ID, names_to = "time_var", values_to = "time_sec") %>%
  mutate(
    ID = as.character(ID),
    # TIME005:TIME025 correspond to experimental positions 1:21.
    position = as.integer(stringr::str_remove(time_var, "TIME")) - 4L,
    time_sec = readr::parse_double(as.character(time_sec))
  ) %>%
  left_join(
    task1 %>%
      transmute(
        ID = as.character(ID),
        position,
        trial,
        vistype
      ),
    by = c("ID", "position")
  ) %>%
  filter(!is.na(time_sec), !is.na(vistype))

task1_time |>
  slice_head(n = 20) |>
  mutate(vistype = table_vistype(vistype)) |>
  knitr::kable(
    format = "html",
    digits = 2,
    col.names = c(
      "Participant ID",
      "Timing variable",
      "Time (seconds)",
      "Presentation position",
      "Trial",
      "Visualization type"
    ),
    caption = "Diagnostic preview: first 20 comparison-task completion-time records",
    table.attr = kable_table_attributes
  )
```

```{r}
time_table <- task1_time %>%
  remove_group_iqr_outliers(time_sec, vistype) %>%
  group_by(vistype) %>%
  summarise(
    n = n(),
    median_sec = median(time_sec, na.rm = TRUE),
    mean_sec = mean(time_sec, na.rm = TRUE),
    sd_sec = sd(time_sec, na.rm = TRUE),
    .groups = "drop"
  )

time_table |>
  arrange(median_sec) |>
  mutate(
    vistype = table_vistype(vistype),
    across(c(median_sec, mean_sec, sd_sec), \(x) round(x, 1))
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Visualization type",
      "Responses",
      "Median time (seconds)",
      "Mean time (seconds)",
      "SD (seconds)"
    ),
    caption = "Observed comparison-task completion time by visualization type",
    table.attr = kable_table_attributes
  )
```
```{r}
task1_time %>%
  remove_group_iqr_outliers(time_sec, vistype) %>%
  ggplot(aes(vistype, time_sec)) + 
  geom_boxplot(outlier.alpha = 0.3) +
  labs(
    title = "Time spent by visualization type",
    x = NULL,
    y = "Time spent on page (seconds)"
  ) +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.x = element_text(angle = 35, hjust = 1)
  )
```

### Model-estimated completion time

```{r task1-time-model}
task1_time_model <- task1_time %>%
  transmute(
    ID = factor(ID),
    vistype = relevel(factor(vistype), ref = "basic"),
    position = as.integer(position),
    time_sec = as.numeric(time_sec)
  ) %>%
  filter(
    !is.na(ID),
    !is.na(vistype),
    !is.na(position),
    !is.na(time_sec),
    time_sec > 0
  ) %>%
  mutate(position_c = position - mean(position))

task1_time_fit <- lmer(
  log(time_sec) ~ vistype + position_c + (1 | ID),
  data = task1_time_model
)

time_estimates <- emmeans(
  task1_time_fit,
  ~ vistype,
  type = "response"
) %>%
  confint() %>%
  as.data.frame() %>%
  transmute(
    vistype,
    estimated_time = response,
    lower_ci = lower.CL,
    upper_ci = upper.CL
  ) %>%
  arrange(estimated_time) %>%
  mutate(vistype = factor(vistype, levels = vistype))

time_estimates |>
  mutate(
    vistype = table_vistype(vistype),
    across(
      c(estimated_time, lower_ci, upper_ci),
      \(x) round(x, 1)
    )
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Visualization type",
      "Estimated time (seconds)",
      "Lower 95% CI",
      "Upper 95% CI"
    ),
    caption = "Model-estimated comparison-task completion time",
    table.attr = kable_table_attributes
  )
```

```{r task1-time-estimate-plot}
# Model-estimated time for each visualization type.
ggplot(
  time_estimates,
  aes(x = estimated_time, y = vistype)
) +
  geom_errorbar(
    aes(xmin = lower_ci, xmax = upper_ci),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_point(size = 3.5) +
  geom_text(
    aes(label = sprintf("%.1f s", estimated_time)),
    nudge_y = 0.36,
    vjust = 0.5,
    size = 6
  ) +
  scale_y_discrete(
    expand = expansion(add = 0.5),
    labels = vistype_labels
  ) +
  scale_x_continuous(
    labels = \(x) sprintf("%.0f s", x),
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  labs(
    title = "Comparison task completion time",
    x = "Model-estimated time (seconds)",
    y = NULL,
    caption = paste(
      "Points show back-transformed model estimates;",
      "error bars indicate 95% confidence intervals."
    )
  ) +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.caption = element_text(hjust = 0.5, face = "italic"),
    axis.ticks.y = element_blank()
  )
```
