---
title: "Task 2"
execute:
  warning: false
  message: false
params:
  export_artifacts: false
---

```{r shared-setup}
#| include: false
source("scripts/analysis_setup.R")
```


## Response scoring and trial reconstruction

```{r}
task2_order <- data %>%
  select(ID, A207_01:A207_21) %>%
  pivot_longer(
    -ID,
    names_to = "trial",
    names_prefix = "A207_",
    values_to = "position"
  ) %>%
  mutate(
    ID = as.character(ID),
    trial = as.integer(trial),
    position = as.integer(position)
  )
```

```{r}
task2_design <- data %>%
  select(ID, A205_01) %>%
  tidyr::separate_rows(A205_01, sep = "###") %>%
  filter(A205_01 != "") %>%
  transmute(
    ID,
    trial   = as.integer(get_field(A205_01, "trial")),
    vistype = get_field(A205_01, "vistype"),
    pattern = get_field(A205_01, "pattern"),
    subject = as.integer(get_field(A205_01, "subject")),
    stimulus = get_field(A205_01, "stim"),
    img_url  = get_field(A205_01, "imgUrl"),
    input_id = get_field(A205_01, "inputId")
  )

task2_responses <- data %>%
  select(ID, A201_01:A201_21) %>%
  pivot_longer(
    -ID,
    names_to = "input_id",
    values_to = "selected_region"
  )
```


```{r}
score_text <- paste(
  readLines(project_path("regions.txt"), encoding = "UTF-8"),
  collapse = "\n"
)
score_matches <- stringr::str_match_all(
  score_text,
  '"([^"]+)"\\s*=>\\s*array\\(([^)]*)\\)'
)[[1]]
region_scores <- tibble::tibble(
  stimulus = score_matches[, 2],
  values = score_matches[, 3]
) %>%
  tidyr::separate_rows(values, sep = ",\\s*") %>%
  group_by(stimulus) %>%
  mutate(
    region = sprintf("R%02d", row_number()),
    score = as.numeric(values)
  ) %>%
  ungroup() %>%
  select(stimulus, region, score)
```

```{r}
task2 <- task2_design %>%
  mutate(ID = as.character(ID)) %>%
  left_join(task2_responses %>% mutate(ID = as.character(ID)),
            by = c("ID", "input_id")) %>%
  mutate(
    region_number = readr::parse_number(selected_region),
    region = sprintf("R%02d", region_number)
  ) %>%
  left_join(region_scores, by = c("stimulus", "region")) %>%
  left_join(task2_order, by = c("ID", "trial")) %>%
  select(ID, trial, position, vistype, stimulus, region, score)

task2 |>
  slice_head(n = 20) |>
  mutate(vistype = table_vistype(vistype)) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Participant ID",
      "Trial",
      "Presentation position",
      "Visualization type",
      "Stimulus",
      "Selected region",
      "Score"
    ),
    caption = "Diagnostic preview: first 20 reconstructed detection-task responses",
    table.attr = kable_table_attributes
  )
```




## Descriptive detection scores

```{r}
task2_model <- task2 %>%
  transmute(
    ID = factor(ID),
    stimulus = factor(stimulus),
    vistype = relevel(factor(vistype), ref = "basic"),
    position_c = as.integer(position) - 11,
    score = as.numeric(score)
  ) %>%
  filter(complete.cases(.))
task2_model |>
  slice_head(n = 20) |>
  mutate(vistype = table_vistype(vistype)) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Participant ID",
      "Stimulus",
      "Visualization type",
      "Centered presentation position",
      "Score"
    ),
    caption = "Diagnostic preview: first 20 detection-task model records",
    table.attr = kable_table_attributes
  )
```

```{r}
task2_model |>
  filter(stimulus == "MMN_30") |>
  slice_head(n = 20) |>
  mutate(vistype = table_vistype(vistype)) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Participant ID",
      "Stimulus",
      "Visualization type",
      "Centered presentation position",
      "Score"
    ),
    caption = "Diagnostic preview: first 20 model records for stimulus MMN_30",
    table.attr = kable_table_attributes
  )
```

```{r task2-descriptive-summary}
task2_table <- task2_model %>%
  group_by(vistype) %>%
  summarise(
    n = n(),
    mean = round(mean(score), 3),
    median = round(median(score), 3),
    sd = round(sd(score), 3),
    .groups = "drop"
  )

task2_table |>
  arrange(desc(median)) |>
  mutate(vistype = table_vistype(vistype)) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Visualization type",
      "Responses",
      "Mean score",
      "Median score",
      "SD"
    ),
    caption = "Observed detection-task scores by visualization type",
    table.attr = kable_table_attributes
  )
```

```{r task2-descriptive-plot}
task2_table |>
  ggplot(aes(median, reorder(vistype, median))) +
  geom_col(width = 0.7, fill = "#6BAED6") +
  geom_errorbar(
    aes(
      xmin = pmax(median - sd, 0),
      xmax = pmin(median + sd, 1)
    ),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_text(
    aes(label = sprintf("%.2f", median)),
    hjust = -0.15,
    nudge_y = -0.20,
    size = 6
  ) +
  scale_x_continuous(
    "Median score",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.10))
  ) +
  scale_y_discrete(labels = vistype_labels) +
  labs(title = "Detection task accuracy", y = NULL) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 22),
    axis.text = element_text(size = 20),
    axis.ticks.y = element_blank()
  )
```

## Detection-score analysis

```{r}
model_task2 <- glmmTMB(
  score ~ vistype + 
    (1 | ID),
  family = ordbeta(link = "logit"),
  data = task2_model
)

summary(model_task2)$coefficients$cond |>
  as.data.frame() |>
  tibble::rownames_to_column("term") |>
  transmute(
    term = dplyr::recode(
      term,
      `(Intercept)` = "Basic topoplot (intercept)",
      vistypeadjacent = "Adjacent plots vs Basic topoplot",
      vistypeanimation = "Animation vs Basic topoplot",
      vistypebivariate_corner = "Bivariate corner vs Basic topoplot",
      vistypebivariate_range = "Bivariate range vs Basic topoplot",
      vistypetriple_CI = "Confidence intervals vs Basic topoplot",
      vistypeuncert_markers = "Uncertainty markers vs Basic topoplot",
      .default = term
    ),
    estimate = Estimate,
    standard_error = .data[["Std. Error"]],
    z_value = .data[["z value"]],
    p_value = format.pval(
      .data[["Pr(>|z|)"]],
      digits = 3,
      eps = 0.001
    )
  ) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Term",
      "Estimate (logit scale)",
      "Standard error",
      "z value",
      "p value"
    ),
    caption = "Fixed effects for the detection-score model",
    table.attr = kable_table_attributes
  )
```

```{r}
model_task2_no_vistype <- update(
  model_task2,
  . ~ . - vistype
)

anova(model_task2_no_vistype, model_task2) |>
  as.data.frame() |>
  tibble::rownames_to_column("model") |>
  mutate(
    model = dplyr::recode(
      model,
      model_task2_no_vistype = "Intercept only",
      model_task2 = "Visualization type",
      .default = model
    ),
    `Pr(>Chisq)` = format.pval(
      .data[["Pr(>Chisq)"]],
      digits = 3,
      eps = 0.001
    )
  ) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Model",
      "Parameters",
      "AIC",
      "BIC",
      "Log likelihood",
      "Deviance",
      "Likelihood-ratio chi-square",
      "df difference",
      "p value"
    ),
    caption = "Likelihood-ratio test of visualization type for detection score",
    table.attr = kable_table_attributes
  )
```

```{r task2-estimates}
task2_estimates <- emmeans(
  model_task2,
  ~ vistype,
  type = "response"
) |>
  confint() |>
  as.data.frame() |>
  transmute(
    vistype,
    predicted_score = response,
    lower_ci = asymp.LCL,
    upper_ci = asymp.UCL
  ) |>
  arrange(predicted_score) |>
  mutate(
    vistype = factor(vistype, levels = vistype)
  )

task2_estimates |>
  mutate(
    vistype = table_vistype(vistype),
    across(
      c(predicted_score, lower_ci, upper_ci),
      \(x) scales::percent(x, accuracy = 0.1)
    )
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Visualization type",
      "Estimated detection score",
      "Lower 95% CI",
      "Upper 95% CI"
    ),
    caption = "Model-estimated detection scores by visualization type",
    table.attr = kable_table_attributes
  )
```
```{r task2-pairwise-comparisons}
emm_task2 <- emmeans(
  model_task2,
  ~ vistype
)

task2_pairs <- pairs(
  emm_task2,
  adjust = "tukey"
)

task2_pairs |>
  as.data.frame() |>
  mutate(contrast = table_contrast(contrast)) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Comparison",
      "Difference (logit scale)",
      "SE",
      "df",
      "z ratio",
      "Tukey-adjusted p value"
    ),
    caption = "Tukey-adjusted pairwise comparisons of detection scores",
    table.attr = kable_table_attributes
  )

task2_pairs |>
  as.data.frame() |>
  dplyr::filter(p.value < 0.05) |>
  mutate(contrast = table_contrast(contrast)) |>
  knitr::kable(
    format = "html",
    digits = 3,
    col.names = c(
      "Comparison",
      "Difference (logit scale)",
      "SE",
      "df",
      "z ratio",
      "Tukey-adjusted p value"
    ),
    caption = "Statistically significant Tukey-adjusted pairwise comparisons",
    table.attr = kable_table_attributes
  )
```
```{r task2-response-scale-comparisons}
emm_task2_response <- regrid(
  emm_task2,
  transform = "response"
)

task2_score_differences <- pairs(
  emm_task2_response,
  adjust = "tukey"
) |>
  confint() |>
  as.data.frame()

task2_score_differences |>
  mutate(
    contrast = table_contrast(contrast),
    across(where(is.numeric), \(x) round(x, 3))
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Contrast",
      "Difference",
      "SE",
      "df",
      "Lower 95% CI",
      "Upper 95% CI"
    ),
    caption = "Response-scale pairwise differences in detection scores",
    table.attr = kable_table_attributes
  )
```


```{r}

detection_axis_colours <- ifelse(
  levels(task2_estimates$vistype) == "basic",
  "#377EB8",
  "#222222"
)

detection_score_figure <- ggplot(
  task2_estimates,
  aes(
    x = predicted_score,
    y = vistype,
    colour = vistype == "basic"
  )
) +
  geom_errorbar(
    aes(
      xmin = lower_ci,
      xmax = upper_ci
    ),
    orientation = "y",
    width = 0.14,
    linewidth = 0.55
  ) +
  geom_point(size = 3) +
  geom_text(
    aes(label = scales::percent(predicted_score, accuracy = 1)),
    nudge_y = 0.36,
    vjust = 0.5,
    size = 5.5
  ) +
  scale_y_discrete(
    labels = function(x) {
      publication_labels <- c(
        basic = "Basic topoplot",
        adjacent = "Adjacent plots",
        bivariate_corner = "Bivariate corners",
        bivariate_range = "Bivariate range",
        triple_CI = "Confidence intervals",
        uncert_markers = "Uncertainty markers",
        animation = "Animation"
      )
      unname(publication_labels[x])
    },
    expand = expansion(add = 0.5)
  ) +
  scale_colour_manual(
    values = c(`FALSE` = "#222222", `TRUE` = "#377EB8"),
    guide = "none"
  ) +
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    limits = c(0.60, 0.75),
    breaks = seq(0.60, 0.75, 0.05),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    title = "Detection task score",
    subtitle = "Model-estimated detection score with 95% confidence intervals",
    x = "Model-estimated detection score",
    y = NULL
  ) +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 22
    ),
    plot.title.position = "plot",
    plot.subtitle = element_text(hjust = 0.5, size = 18),
    axis.text.x = element_text(size = 20),
    axis.text.y = element_text(
      size = 24,
      colour = detection_axis_colours,
      margin = margin(r = 12)
    ),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

detection_score_figure

ggsave(
  project_path("figures", "detection_score_figure.png"),
  detection_score_figure,
  width = 9,
  height = 5.5,
  dpi = 180
)
```

### Detection-score differences relative to Basic topoplot

This plot shows response-scale differences from Basic topoplot. Confidence intervals crossing zero indicate that the corresponding visualization is not clearly different from Basic at the 95% confidence level.

```{r task2-basic-contrast-figure}
task2_basic_reference <- which(
  emm_task2_response@levels$vistype == "basic"
)

task2_basic_contrasts_raw <- contrast(
  emm_task2_response,
  method = "trt.vs.ctrl",
  ref = task2_basic_reference
) |>
  confint() |>
  as.data.frame()

task2_lower_ci_column <- grep(
  "lower.CL$|asymp.LCL$",
  names(task2_basic_contrasts_raw),
  value = TRUE
)[1]
task2_upper_ci_column <- grep(
  "upper.CL$|asymp.UCL$",
  names(task2_basic_contrasts_raw),
  value = TRUE
)[1]

task2_basic_contrasts <- task2_basic_contrasts_raw |>
  transmute(
    vistype = stringr::str_remove(contrast, " - basic$"),
    difference = estimate,
    lower_ci = .data[[task2_lower_ci_column]],
    upper_ci = .data[[task2_upper_ci_column]]
  ) |>
  mutate(
    vistype = factor(vistype, levels = levels(task2_estimates$vistype))
  )

task2_contrast_limit <- max(
  abs(c(task2_basic_contrasts$lower_ci, task2_basic_contrasts$upper_ci)),
  na.rm = TRUE
)
task2_contrast_limit <- max(
  0.05,
  ceiling(task2_contrast_limit / 0.05) * 0.05
)

task2_contrast_labels <- c(
  adjacent = "Adjacent plots",
  bivariate_corner = "Bivariate corners",
  bivariate_range = "Bivariate range",
  triple_CI = "Confidence intervals",
  uncert_markers = "Uncertainty markers",
  animation = "Animation"
)

detection_score_contrast_figure <- ggplot(
  task2_basic_contrasts,
  aes(x = difference, y = vistype)
) +
  geom_vline(
    xintercept = 0,
    linetype = "dashed",
    linewidth = 0.6,
    colour = "#377EB8"
  ) +
  geom_errorbar(
    aes(xmin = lower_ci, xmax = upper_ci),
    orientation = "y",
    width = 0.14,
    linewidth = 0.55,
    colour = "#333333"
  ) +
  geom_point(size = 3, colour = "#333333") +
  geom_text(
    aes(label = ifelse(
      abs(difference) < 0.005,
      "0 pp",
      sprintf("%+.0f pp", 100 * difference)
    )),
    nudge_y = 0.32,
    vjust = 0.5,
    size = 5.5,
    colour = "#333333"
  ) +
  scale_y_discrete(
    labels = function(x) unname(task2_contrast_labels[x]),
    expand = expansion(add = 0.5)
  ) +
  scale_x_continuous(
    limits = c(-task2_contrast_limit, task2_contrast_limit),
    labels = function(x) ifelse(
      abs(x) < 0.005,
      "0 pp",
      sprintf("%+.0f pp", 100 * x)
    ),
    breaks = scales::breaks_pretty(n = 5),
    expand = expansion(mult = 0.04)
  ) +
  labs(
    title = "Detection task score differences",
    subtitle = "Response-scale contrasts with 95% confidence intervals",
    x = NULL,
    y = NULL
  ) +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 22),
    plot.subtitle = element_text(hjust = 0.5, size = 18),
    plot.title.position = "plot",
    axis.text.y = element_text(size = 24, margin = margin(r = 12)),
    axis.text.x = element_text(size = 20),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank()
  )

detection_score_contrast_caption <- ggplot() +
  annotate(
    "text",
    x = 0.497,
    y = 0.5,
    label = "Difference in estimated detection score vs",
    hjust = 1,
    size = 20 / ggplot2::.pt,
    family = "Lato",
    colour = "#222222"
  ) +
  annotate(
    "text",
    x = 0.503,
    y = 0.5,
    label = "Basic topoplot",
    hjust = 0,
    size = 20 / ggplot2::.pt,
    family = "Lato",
    colour = "#377EB8"
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), clip = "off") +
  theme_void()

detection_score_contrast_figure <-
  detection_score_contrast_figure /
  detection_score_contrast_caption +
  plot_layout(heights = c(1, 0.10))

detection_score_contrast_figure

ggsave(
  project_path("figures", "detection_score_contrast_figure.png"),
  detection_score_contrast_figure,
  width = 9,
  height = 5.25,
  dpi = 180
)
```

<!-- Descriptive code moved before the model analysis. -->
```{r}
#| eval: false
#| include: false
task2_table <- task2_model %>%
  group_by(vistype) %>%
  summarise(
    n = n(),
    mean = round(mean(score), 3),
    median = round(median(score), 3),
    sd = round(sd(score), 3),
    .groups = "drop"
  )

task2_table |> arrange(desc(median))
```
```{r}
#| eval: false
#| include: false
task2_table |>
  ggplot(aes(median, reorder(vistype, median))) +
  geom_col(
    width = 0.7,
    fill = "#6BAED6"
  ) +
  geom_errorbar(
    aes(
      xmin = pmax(median - sd, 0),
      xmax = pmin(median + sd, 1)
    ),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_text(
    aes(label = sprintf("%.2f", median)),
    hjust = -0.15,
    nudge_y = -0.20,
    size = 6
  ) +
  scale_x_continuous(
    "Median score",
    limits = c(0, 1),
    breaks = seq(0, 1, 0.2),
    expand = expansion(mult = c(0, 0.10))
  ) +
  scale_y_discrete(
    labels = c(
      adjacent = "Adjacent\nplots",
      animation = "Animation",
      bivariate_corner = "Bivariate\ncorner",
      bivariate_range = "Bivariate\nrange",
      triple_CI = "Confidence\nintervals",
      uncert_markers = "Uncertainty\nmarkers",
      basic = "Basic\ntopoplot"
    )
  ) +
  labs(
    title = "Detection task accuracy",
    y = NULL
  ) +
  coord_cartesian(clip = "off") +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(
      hjust = 0.5,
      face = "bold",
      size = 22
    ),
    axis.text = element_text(size = 20),
    axis.ticks.y = element_blank()
  )
```


## Completion-time analysis

```{r task2-time-data}
task2_time <- data %>%
  select(ID, TIME029:TIME049) %>%
  pivot_longer(
    -ID,
    names_to = "time_var",
    values_to = "time_sec"
  ) %>%
  mutate(
    ID = as.character(ID),
    # TIME029:TIME049 correspond to experimental positions 1:21.
    position = as.integer(stringr::str_remove(time_var, "TIME")) - 28L,
    time_sec = readr::parse_double(as.character(time_sec))
  ) %>%
  left_join(
    task2 %>%
      transmute(
        ID = as.character(ID),
        position,
        trial,
        vistype
      ),
    by = c("ID", "position")
  ) %>%
  filter(!is.na(time_sec), !is.na(vistype), time_sec > 0)

task2_time |>
  slice_head(n = 20) |>
  mutate(vistype = table_vistype(vistype)) |>
  knitr::kable(
    format = "html",
    digits = 2,
    col.names = c(
      "Participant ID",
      "Timing variable",
      "Time (seconds)",
      "Presentation position",
      "Trial",
      "Visualization type"
    ),
    caption = "Diagnostic preview: first 20 detection-task completion-time records",
    table.attr = kable_table_attributes
  )
```

```{r task2-time-summary}
task2_time_table <- task2_time %>%
  remove_group_iqr_outliers(time_sec, vistype) %>%
  group_by(vistype) %>%
  summarise(
    n = n(),
    median_sec = median(time_sec, na.rm = TRUE),
    mean_sec = mean(time_sec, na.rm = TRUE),
    sd_sec = sd(time_sec, na.rm = TRUE),
    .groups = "drop"
  )

task2_time_table |>
  arrange(median_sec) |>
  mutate(
    vistype = table_vistype(vistype),
    across(c(median_sec, mean_sec, sd_sec), \(x) round(x, 1))
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Visualization type",
      "Responses",
      "Median time (seconds)",
      "Mean time (seconds)",
      "SD (seconds)"
    ),
    caption = "Observed detection-task completion time by visualization type",
    table.attr = kable_table_attributes
  )
```

### Model-estimated completion time

```{r task2-time-model}
task2_time_model <- task2_time %>%
  transmute(
    ID = factor(ID),
    vistype = relevel(factor(vistype), ref = "basic"),
    position = as.integer(position),
    time_sec = as.numeric(time_sec)
  ) %>%
  filter(complete.cases(.), time_sec > 0) %>%
  mutate(position_c = position - mean(position))

task2_time_fit <- lmer(
  log(time_sec) ~ vistype + position_c + (1 | ID),
  data = task2_time_model
)

task2_time_estimates <- emmeans(
  task2_time_fit,
  ~ vistype,
  type = "response"
) %>%
  confint() %>%
  as.data.frame() %>%
  transmute(
    vistype,
    estimated_time = response,
    lower_ci = lower.CL,
    upper_ci = upper.CL
  ) %>%
  arrange(estimated_time) %>%
  mutate(vistype = factor(vistype, levels = vistype))

task2_time_estimates |>
  mutate(
    vistype = table_vistype(vistype),
    across(
      c(estimated_time, lower_ci, upper_ci),
      \(x) round(x, 1)
    )
  ) |>
  knitr::kable(
    format = "html",
    col.names = c(
      "Visualization type",
      "Estimated time (seconds)",
      "Lower 95% CI",
      "Upper 95% CI"
    ),
    caption = "Model-estimated detection-task completion time",
    table.attr = kable_table_attributes
  )
```

```{r task2-time-estimate-plot}
#| fig-width: 9
#| fig-height: 5.5
ggplot(
  task2_time_estimates,
  aes(x = estimated_time, y = vistype)
) +
  geom_errorbar(
    aes(xmin = lower_ci, xmax = upper_ci),
    orientation = "y",
    width = 0.18,
    linewidth = 0.7
  ) +
  geom_point(size = 3.5) +
  geom_text(
    aes(label = sprintf("%.1f s", estimated_time)),
    nudge_y = 0.36,
    vjust = 0.5,
    size = 6
  ) +
  scale_y_discrete(
    expand = expansion(add = 0.5),
    labels = vistype_labels
  ) +
  scale_x_continuous(
    labels = \(x) sprintf("%.0f s", x),
    expand = expansion(mult = c(0.05, 0.12))
  ) +
  labs(
    title = "Detection task completion time",
    x = "Model-estimated time (seconds)",
    y = NULL,
    caption = paste0(
      "Points show back-transformed model estimates.\n",
      "Error bars indicate 95% confidence intervals."
    )
  ) +
  theme_classic(base_family = "Lato", base_size = 20) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.caption = element_text(hjust = 0.5, face = "italic"),
    axis.ticks.y = element_blank(),
    plot.margin = margin(8, 16, 12, 16)
  )
```
